---
title: "Trophic transfer of polyunsaturated fatty acids across the aquatic-terrestrial interface: an experimental tritrophic food chain approach: Statistics"
output:
  html_document:
    figure_caption: yes
    highlight: tango
    number_sections: yes
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

# Required packages
```{r, message=FALSE, warning=FALSE}
library(data.table) # Data wrangling
library(dplyr) # Data wrangling
library(tidyr) # Data wrangling
library(tidylog, warn.conflicts = FALSE) # Data wrangling
library(plyr) # Data wrangling

library(ggplot2) # plots
library(ggpubr) # plots
library(ggrepel) # plots

library(smatr) # standardized major axis regression

library(vegan) # NMDS, ANOSIM

library(glmmTMB) # LMM
library(DAAG) # Model diagnostics

library(effects) # plot interactions

library(car) # type II ANOVA

library(sjPlot) # Get summary table of LMM and LM
library(texreg) # export model output to html or latex

library(lubridate) # handle date and time data
```

# Set path
```{r, message=FALSE, warning=FALSE}
path <-"/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics"
setwd(path)
```

# Load data
```{r, message=FALSE, warning=FALSE}
spider_treatment <- read.table("/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/response_spiders/spider_to_treatment_3.csv", 
                               header = TRUE, 
                               sep = ",",na.strings = c("NR", "NA"), 
                               stringsAsFactors = FALSE, strip.white = TRUE) 

microcosm <- read.table("/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/PUFA_data/microcosm_pufa.csv", header = TRUE, 
                       sep = ",", na.strings = c("NR", "NA"), 
                       stringsAsFactors = FALSE, strip.white = TRUE) 

encapsulation <- read.table("/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/response_spiders/encapsulation_assay_measurements_6.csv", header = TRUE, 
                            sep = ",",na.strings = c("NR", "NA"), 
                            stringsAsFactors = FALSE, strip.white = TRUE) 

encapsulation_form <- read.table("/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/response_spiders/encapsulation_assay_form_sheet_2.csv", header = TRUE,
                                 sep = ",",na.strings = c("NR", "NA"),
                                 stringsAsFactors = FALSE, strip.white = TRUE) 

condition <- read.table("/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/response_spiders/fitness_spiders_5.csv", header = TRUE, 
                        sep = ",",na.strings = c("NR", "NA"), 
                        stringsAsFactors = FALSE, strip.white = TRUE) 

growth <- read.table("/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/response_spiders/growth_spiders_190722_5.csv", header = TRUE, 
                     sep = ",",na.strings = c("NR", "NA"), 
                     stringsAsFactors = FALSE, strip.white = TRUE) 
```

# Prepare data for merging
## Encapsulation data

Merge data tables of encapsulation assay
```{r, message=FALSE, warning=FALSE}
encapsulation <- full_join(encapsulation, encapsulation_form, 
                       by = c("Organism", "treatment", "experiment", 
                              "experiment_ID", "spider_ID",
                              "spider_ID_experiment_ID"))
```

Calculate area of not encapsulated monofilament 

$$ monofilament_{area} = A_{not \  encapsulated} = length \cdot width $$
```{r, message=FALSE, warning=FALSE}
encapsulation$area_monofilament <- with(encapsulation, length_mm*width_mm)
```

Calculate Ratio of encapsulated area and area of not encapsulated monofilament 

$$ R_{e/ne} = {A_{encapsulated}  \over A_{not \ encapsulated}}$$
```{r, message=FALSE, warning=FALSE}
encapsulation$R_e_ne <- with(encapsulation, encapsulated_area_mm_2/area_monofilament)
```

Calculate area of encapsulation

$$ A_{encapsulation} = {A_{encapsulated}  - A_{not \ encapsulated}}$$
```{r, message=FALSE, warning=FALSE}
encapsulation$area_encapsulation <- with(encapsulation,
                                         encapsulated_area_mm_2-area_monofilament)
```

If area encapsulation < 0: Set area encapsulated = 0, because there cannot be a negative area encapsulation
```{r, message=FALSE, warning=FALSE}
encapsulation$area_encapsulation <- with(encapsulation, ifelse(area_encapsulation > 0, area_encapsulation, 0))
```

For statistics: Calculate mean of measurements per treatment and date 
```{r, message=FALSE, warning=FALSE}
encapsulation_stat <- ddply(encapsulation, c("Organism", "treatment", "experiment", 
                            "experiment_ID","spider_ID", "spider_ID_experiment_ID",
                            "start_encapsulation"),
                            summarise, 
                            mean_R_e_ne = mean(R_e_ne),
                            mean_area_encapsulation = mean(area_encapsulation))
```

Remove test spiders
```{r, message=FALSE, warning=FALSE}
encapsulation_stat <- filter(encapsulation_stat, experiment != "")
```

Check NAs
```{r, message=FALSE, warning=FALSE}
filter(encapsulation_stat, is.na(mean_R_e_ne))
filter(encapsulation_stat, is.na(mean_area_encapsulation))
```

NAs right, because monofilament not found after 24 h encapsulation

Rename start encapsulation
```{r, message=FALSE, warning=FALSE}
names(encapsulation_stat)[7] <- "date_endpoint"
```

date in date format
```{r, message=FALSE, warning=FALSE}
encapsulation_stat$date_endpoint <- as.Date(encapsulation_stat$date_endpoint,
                                   "%d.%m.%Y")
```

## Growth data

Get variables for statistics
```{r, message=FALSE, warning=FALSE}
growth_stat <- select(growth, c("date", "Organism", "experiment",
                            "experiment_ID", "spider_ID", "treatment",
                            "weight_spider_g"))
```

Correct date for the following spiders:
* 22.05.2019 fishfood_2 95: died at 14.05.2019, frozen since death and weighed at 22.05.2019
* 22.05.2019 fishfood_13 121: died at 14.05.2019, frozen since death and weighed at 22.05.2019
* 22.05.2019 fishfood_16 108: died at 17.05.2019, frozen since death and weighed at 22.05.2019
* 22.05.2019 algae_4 88: died at 03.05.2019, frozen since death and weighed at 22.05.2019
* 22.05.2019 algae_19 105: died at 07.05.2019, frozen since death and weighed at 22.05.2019

```{r, message=FALSE, warning=FALSE}
growth_stat$date[growth_stat$experiment_ID == "fishfood_2" &
                   growth_stat$date == "22.05.2019"]  <- "14.05.2019"
growth_stat$date[growth_stat$experiment_ID == "fishfood_13" 
                 & growth_stat$date == "22.05.2019"]  <- "14.05.2019"  
growth_stat$date[growth_stat$experiment_ID == "fishfood_16" 
                 & growth_stat$date == "22.05.2019"]  <- "17.05.2019"  
growth_stat$date[growth_stat$spider_ID == "88" 
                 & growth_stat$date == "22.05.2019"]  <- "03.05.2019"  
growth_stat$date[growth_stat$spider_ID == "105" 
                 & growth_stat$date == "22.05.2019"]  <- "07.05.2019" 
```

Remove weight of spiders, that were weighed 2 times:
* 22.05.2019 oatmeal_15 99: died at 15.05.2019 and also weighed there, but frozen since death and weighed again at 22.05.2019, so not use weight from 22.05.2019
* 22.05.2019 algae_18: died at 08.05.2019, and also weighed there, but frozen since death and weighed again at 22.05.2019, so not use weight from 22.05.2019

* 15.07.2019 oatmeal_6 167: weighed one time before and one time after feeding. So not use weight after feeding: 0.0194
```{r, message=FALSE, warning=FALSE}
growth_stat <- filter(growth_stat, experiment_ID != "algae_18" |
                        date != "22.05.2019")

growth_stat <- filter(growth_stat, experiment_ID != "oatmeal_15" |
                        date != "22.05.2019")

growth_stat <- filter(growth_stat, experiment_ID != "oatmeal_6" |
                        weight_spider_g != "0.0194")

```

date in date format
```{r, message=FALSE, warning=FALSE}
growth_stat$date <- as.Date(growth_stat$date, "%d.%m.%Y")
names(growth_stat)[1] <- "date_endpoint"
```

Remove "leaves" spiders from M1: Treatment was ended, because not enough leaves chiros to feed spiders
```{r, message=FALSE, warning=FALSE}
growth_stat <- filter(growth_stat, experiment != "M1" | treatment != "leaves")
```

## Condition data

Remove 15.07.2019 oatmeal_6 167: picture one time before and one time after feeding. So not use picture after feeding
```{r, message=FALSE, warning=FALSE}
condition <- filter(condition, comment != "foto nach fütterung")
```

Calculate body condition

$$ body\ condition = { prosoma\  width \over opisthosoma\  width} = { thorax\  width \over abdomen\  width} $$
```{r, message=FALSE, warning=FALSE}
condition$body_cond <- with(condition, prosoma_width_cm/opisthosoma_width_cm)
```

Choose variables for statistics
```{r, message=FALSE, warning=FALSE}
condition_stat <- select(condition, -c("picture_ID", "prosoma_length_cm", "opisthosoma_length_cm", "comment"))
```

Correct date for the following spiders:
* 22.05.2019 fishfood_2 95: died at 14.05.2019, frozen since death and picture at 22.05.2019
* 22.05.2019 fishfood_13 121: died at 14.05.2019, frozen since death and picture at 22.05.2019
* 22.05.2019 fishfood_16 108: died at 17.05.2019, frozen since death and picture at 22.05.2019
* 22.05.2019 algae_4 88: died at 03.05.2019, frozen since death and weighed at 22.05.2019

* 26.07.2019 oatmeal_13	189: in encapsulation assay at 22.07.2019, but picture taken on 26.07.2019 	

```{r, message=FALSE, warning=FALSE}
condition_stat$date[condition_stat$experiment_ID == "fishfood_2" &
                   condition_stat$date == "22.05.2019"]  <- "14.05.2019"
condition_stat$date[condition_stat$experiment_ID == "fishfood_13" 
                 & condition_stat$date == "22.05.2019"]  <- "14.05.2019"  
condition_stat$date[condition_stat$experiment_ID == "fishfood_16" 
                 & condition_stat$date == "22.05.2019"]  <- "17.05.2019" 
condition_stat$date[condition_stat$spider_ID == "88" 
                 & condition_stat$date == "22.05.2019"]  <- "03.05.2019"

condition_stat$date[condition_stat$experiment_ID == "oatmeal_13" 
                 & condition_stat$date == "26.07.2019"]  <- "22.07.2019" 
```

Remove spiders with 2 measurements:

* 22.05.2019 spider algae M1 algae_18 92: spider died at 08.05.2019, and also picture taken there, but frozen since death and picture again at 22.05.2019, so not use picture from 22.05.2019
* 22.05.2019 oatmeal_15 99: died at 15.05.2019 and also picture there, but frozen since death and picture again at 22.05.2019, so not use picture from 22.05.2019
```{r, message=FALSE, warning=FALSE}
condition_stat <- filter(condition_stat, experiment_ID != "algae_18" |
                        date != "22.05.2019")

condition_stat <- filter(condition_stat, experiment_ID != "oatmeal_15" |
                        date != "22.05.2019")
```

date in date format
```{r, message=FALSE, warning=FALSE}
condition_stat$date <- as.Date(condition_stat$date, "%d.%m.%Y")
names(condition_stat)[6] <- "date_endpoint"
```

Remove "leaves" spiders from M1: Treatment was ended, because not enough leaves chiros to feed spiders
```{r, message=FALSE, warning=FALSE}
condition_stat <- filter(condition_stat, experiment != "M1" | treatment != "leaves")
```

## Assign spider to treatment

* deadly endpoints: dry weight, encapsulation and PUFA content
* weekly endpoints: fresh weight, condition

microcosm 1
* t0: 29.04.2019: deadly endpoints and weekly endpoints (only fresh weight and condition)
* t2: 08.05.2019: only weekly endpoints 1
* t3: 15.05.2019: only weekly endpoints 1
* t4: 22.05.2019: deadly endpoints and weekly endpoints
* t5: 29.05.2019: only weekly endpoints 1
* t6: 05.06.2019: only weekly endpoints 1
* t7: 12.06.2019: deadly endpoints and weekly endpoints

* ta: 03.05.2019: start for replacement spiders
* tb: 07.05.2019: start for replacement spiders
* tc: 07.06.2019: algae_15 33: spider died, so no official endpoint date
* td: 14.05.2019: fishfood_2, fishfood_13: spider died, so no official endpoint date
* te: 17.05.2019: fishfood_16 108: spider died, so no official endpoint date

microcosm 2
* t0: 08.07.2019: deadly endpoints and weekly endpoints (only fresh weight and condition)
* t1: 15.07.2019: only weekly endpoints 1
* t2: 22.07.2019: deadly endpoints and weekly endpoints
* t3: 29.07.2019: deadly endpoints and weekly endpoints

* ta: 26.07.2019: oatmeal_2, oatmeal_11, oatmeal_15, fishfood_4, fishfood_12, fishfood_19, leaves_20, leaves_6: spiders died, so no official endpoint date
* tb 19.07.2019: leaves_2: spider died, so no official endpoint date 

Split for M1 and M2
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1 <- filter(spider_treatment, experiment == "M1")
spider_treatment_M2 <- filter(spider_treatment, experiment == "M2")
```

Add date of endpoints 
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1$t0 <- "29.04.2019"
spider_treatment_M1$t2 <- "08.05.2019"
spider_treatment_M1$t3 <- "15.05.2019"
spider_treatment_M1$t4 <- "22.05.2019"
spider_treatment_M1$t5 <- "29.05.2019"
spider_treatment_M1$t6 <- "05.06.2019"
spider_treatment_M1$t7 <- "12.06.2019"

spider_treatment_M1$ta <- "03.05.2019"
spider_treatment_M1$tb <- "07.05.2019"
spider_treatment_M1$tc <- "07.06.2019"
spider_treatment_M1$td <- "14.05.2019"
spider_treatment_M1$te <- "17.05.2019"

spider_treatment_M2$t0 <- "08.07.2019"
spider_treatment_M2$t1 <- "15.07.2019"
spider_treatment_M2$t2 <- "22.07.2019"
spider_treatment_M2$t3 <- "29.07.2019"

spider_treatment_M2$ta <- "26.07.2019"
spider_treatment_M2$tb <- "19.07.2019"
```

Put date of endpoints in long format
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1 <- reshape2::melt(spider_treatment_M1, id.vars = c("Organism", "treatment", "experiment", "experiment_ID", "original_spider_ID", "replace_spider_1", "date_replacement_1", "replace_spider_2", "date_replacement_2", "date_treatment_ended", "dead_not_found_encapsulation", "comment"), value.name = c("date_endpoint"), variable.name = c("time_point"))

spider_treatment_M2 <- reshape2::melt(spider_treatment_M2, id.vars = c("Organism", "treatment", "experiment", "experiment_ID", "original_spider_ID", "replace_spider_1", "date_replacement_1", "replace_spider_2", "date_replacement_2", "date_treatment_ended", "dead_not_found_encapsulation", "comment"), value.name = c("date_endpoint"), variable.name = c("time_point"))
```

Dates in date format
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1$date_replacement_1 <- as.Date(spider_treatment_M1$date_replacement_1, "%d.%m.%Y")
spider_treatment_M1$date_replacement_2 <- as.Date(spider_treatment_M1$date_replacement_2, "%d.%m.%Y")
spider_treatment_M1$date_treatment_ended <- as.Date(spider_treatment_M1$date_treatment_ended, "%d.%m.%Y")
spider_treatment_M1$date_endpoint <- as.Date(spider_treatment_M1$date_endpoint, "%d.%m.%Y")

spider_treatment_M2$date_replacement_1 <- as.Date(spider_treatment_M2$date_replacement_1, "%d.%m.%Y")
spider_treatment_M2$date_replacement_2 <- as.Date(spider_treatment_M2$date_replacement_2, "%d.%m.%Y")
spider_treatment_M2$date_treatment_ended <- as.Date(spider_treatment_M2$date_treatment_ended, "%d.%m.%Y")
spider_treatment_M2$date_endpoint <- as.Date(spider_treatment_M2$date_endpoint, "%d.%m.%Y")
```

Remove dates after treatment ended
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1$remove <- with(spider_treatment_M1, ifelse(date_treatment_ended < date_endpoint, "y", "n"))
spider_treatment_M1 <- filter(spider_treatment_M1, remove == "n")

spider_treatment_M2$remove <- with(spider_treatment_M2, ifelse(date_treatment_ended < date_endpoint, "y", "n"))
spider_treatment_M2 <- filter(spider_treatment_M2, remove == "n")
```

Get replacement spiders
```{r, message=FALSE, warning=FALSE}
replace_spider_M1 <- filter(spider_treatment_M1, !is.na(replace_spider_1))

replace_spider_1_M2 <- filter(spider_treatment_M2, !is.na(replace_spider_1))
replace_spider_2_M2 <- filter(spider_treatment_M2, !is.na(replace_spider_2))
```

Add column spider_ID
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1$spider_ID <- spider_treatment_M1$original_spider_ID
spider_treatment_M2$spider_ID <- spider_treatment_M2$original_spider_ID

replace_spider_M1$spider_ID <- replace_spider_M1$replace_spider_1

replace_spider_1_M2$spider_ID <- replace_spider_1_M2$replace_spider_1
replace_spider_2_M2$spider_ID <- replace_spider_2_M2$replace_spider_2
```

Add column with information, if it is original, replace 1 or replace 2 spider
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1$spider <- "original"
spider_treatment_M2$spider <- "original"

replace_spider_M1$spider <- "replace_1"

replace_spider_1_M2$spider <- "replace_1"
replace_spider_2_M2$spider <- "replace_2"
```

Bind tables together
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1 <- rbind(spider_treatment_M1, replace_spider_M1)

spider_treatment_M2 <- rbind(spider_treatment_M2, replace_spider_1_M2, replace_spider_2_M2)
```

Choose needed columns
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1 <- select(spider_treatment_M1, c("Organism", "treatment", "experiment", "experiment_ID", "dead_not_found_encapsulation", "date_endpoint", "spider_ID", "spider", "date_replacement_1", "date_replacement_2", "date_treatment_ended") )

spider_treatment_M2 <- select(spider_treatment_M2, c("Organism", "treatment", "experiment", "experiment_ID", "dead_not_found_encapsulation", "date_endpoint", "spider_ID", "spider", "date_replacement_1", "date_replacement_2", "date_treatment_ended")) 
```
On 03.05. and 07.05 some spiders replaced. Remove these dates for all original spiders and keep only replacement spiders:

replacement spiders 03.05.2019:
M1	oatmeal_5	  123
M1	oatmeal_1	  126
M1	fishfood_7	127
M1	algae_4	    122
M1	algae_19	  124

replacement spiders 07.05.2019
M1	oatmeal_5	123
M1	oatmeal_1	126
M1	algae_19	124
```{r, message=FALSE, warning=FALSE}
`%notin%` <- Negate(`%in%`)

remove <- c(21, 22, 24, 38, 39, 44, 48, 49, 59, 63, 65, 70, 74, 78, 79, 81, 86, 106, 111, 115, 27, 85, 42, 83, 54, 94, 80, 72, 36, 60, 57, 73, 113, 66, 99, 68, 46, 96, 56, 90, 93, 95, 55, 53, 25, 23, 84, 71, 50, 107, 120, 64, 121, 117, 45, 108, 69, 43, 52, 67, 104, 75, 40, 88, 82, 41, 118, 89, 116, 34, 114, 97, 103, 62, 33, 37, 100, 92, 105, 77)

spider_treatment_M1 <- filter(spider_treatment_M1, date_endpoint != "2019-05-03" | spider_ID %notin% remove)

remove <- c(21, 22, 24, 38, 39, 44, 48, 49, 59, 63, 65, 70, 74, 78, 79, 81, 86, 106, 111, 115, 27, 85, 42, 83, 54, 94, 80, 72, 36, 60, 57, 73, 113, 66, 99, 68, 46, 96, 56, 90, 93, 95, 55, 53, 25, 23, 84, 71, 50, 107, 120, 64, 121, 117, 45, 108, 69, 43, 52, 67, 104, 75, 40, 88, 82, 41, 118, 89, 116, 34, 114, 97, 103, 62, 33, 37, 100, 92, 105, 77, 127, 122)

spider_treatment_M1 <- filter(spider_treatment_M1, date_endpoint != "2019-05-07" | spider_ID %notin% remove)
```

Some spiders found dead not on regular endpoint date. These weighed etc. on one day they were found. So remove these dates from all other spiders.

Spiders died on the 14.05.2019
M1 fishfood_2   95
M1 fishfood_13 121

Spiders died on the 17.05.2019
M1 fishfood_16 108

Spider died on the 07.06.2019
M1 algae_15 33
```{r, message=FALSE, warning=FALSE}
remove <- c(21, 22, 24, 38, 39, 44, 48, 49, 59, 63, 65, 70, 74, 78, 79, 81, 86, 106, 111, 115, 27, 85, 42, 83, 54, 94, 80, 72, 36, 60, 57, 73, 113, 66, 99, 68, 46, 96, 56, 90, 93, 55, 53, 25, 23, 84, 71, 50, 107, 120, 64, 117, 45, 108, 69, 43, 52, 67, 104, 75, 40, 88, 82, 41, 118, 89, 116, 34, 114, 97, 103, 62, 33, 37, 100, 92, 105, 77, 126, 123, 127, 122, 124)

spider_treatment_M1 <- filter(spider_treatment_M1, date_endpoint != "2019-05-14" | spider_ID %notin% remove)

remove <- c(21, 22, 24, 38, 39, 44, 48, 49, 59, 63, 65, 70, 74, 78, 79, 81, 86, 106, 111, 115, 27, 85, 42, 83, 54, 94, 80, 72, 36, 60, 57, 73, 113, 66, 99, 68, 46, 96, 56, 90, 93, 95, 55, 53, 25, 23, 84, 71, 50, 107, 120, 64, 121, 117, 45, 69, 43, 52, 67, 104, 75, 40, 88, 82, 41, 118, 89, 116, 34, 114, 97, 103, 62, 33, 37, 100, 92, 105, 77, 126, 123, 127, 122, 124)

spider_treatment_M1 <- filter(spider_treatment_M1, date_endpoint != "2019-05-17" | spider_ID %notin% remove)

remove <- c(21, 22, 24, 38, 39, 44, 48, 49, 59, 63, 65, 70, 74, 78, 79, 81, 86, 106, 111, 115, 27, 85, 42, 83, 54, 94, 80, 72, 36, 60, 57, 73, 113, 66, 99, 68, 46, 96, 56, 90, 93, 95, 55, 53, 25, 23, 84, 71, 50, 107, 120, 64, 121, 117, 45, 108, 69, 43, 52, 67, 104, 75, 40, 88, 82, 41, 118, 89, 116, 34, 114, 97, 103, 62,
37, 100, 92, 105, 77, 126, 123, 127, 122, 124)

spider_treatment_M1 <- filter(spider_treatment_M1, date_endpoint != "2019-06-07" | spider_ID %notin% remove)
```

No endpoints could be taken, because spiders were not found or only few legs remained
05.06.2019
M1 algae_7 118 (not found)
M1 algae_15 33 (not found)
M1 fishfood_9 50 (only few legs remained)
M1 oatmeal_17 46 (not found)

22.05.2019
M1 oatmeal_14 66 (not found)

29.05.2019
M1 oatmeal_13 113 (not found)
M1 oatmeal_3 42 (not found)
```{r, message=FALSE, warning=FALSE}
remove <- c(118, 33, 50, 46)

spider_treatment_M1 <- filter(spider_treatment_M1, date_endpoint != "2019-06-05" | spider_ID %notin% remove)

remove <- c(66)

spider_treatment_M1 <- filter(spider_treatment_M1, date_endpoint != "2019-05-22" | spider_ID %notin% remove)

remove <- c(42, 113)

spider_treatment_M1 <- filter(spider_treatment_M1, date_endpoint != "2019-05-29" | spider_ID %notin% remove)
```

Some spiders found dead not on regular endpoint date. These weighed and so on on day they were found. So remove these dates from all other spiders.

Spiders died on the 26.07.2019:
M2	oatmeal_11  	130
M2	oatmeal_15    134
M2	oatmeal_2	    218
M2	leaves_20	    208
M2	leaves_6	    226
M2	fishfood_4	  136
M2	fishfood_12	  166
M2	fishfood_19	  177

Spiders died on the 19.07.2019:
M2  leaves_2  197

```{r, message=FALSE, warning=FALSE}
remove <- c(133, 138, 163, 164, 165, 168, 181, 187, 202, 203, 215, 216, 225, 221, 171, 186, 159, 204, 167, 222, 132, 146, 190, 154, 189, 195, 223, 157, 228, 135, 160, 198, 170, 219, 175, 196, 180, 161, 213, 158, 206, 139, 214, 201, 131, 212, 140, 185, 137, 199, 188, 207, 197, 169, 150, 152, 149, 142, 191, 174, 179, 193, 220, 227, 183, 194, 144, 176, 210, 192, 217, 148, 147, 153, 184, 209, 211, 178, 172, 182, 162)

spider_treatment_M2 <- filter(spider_treatment_M2, date_endpoint != "2019-07-26" | spider_ID %notin% remove)

remove <- c(133, 138, 163, 164, 165, 168, 181, 187, 202, 203, 215, 216, 225, 221, 171, 186, 159, 204, 167, 222, 132, 146, 190, 130, 154, 189, 195, 134, 223, 157, 228, 135, 160, 198, 170, 219, 175, 196, 180, 161, 213, 158, 206, 139, 214, 201, 131, 212, 140, 185, 137, 199, 188, 207, 169, 150, 152, 226, 149, 142, 191, 174, 179, 193, 220, 227, 183, 194, 144, 176, 210, 192, 218, 217, 148, 147, 136, 166, 177, 153, 184, 209, 211, 178, 172, 208, 182, 162)

spider_treatment_M2 <- filter(spider_treatment_M2, date_endpoint != "2019-07-19" | spider_ID %notin% remove)
```

No endpoints could be taken, because spiders were not found 

29.07.2019
M2 fishfood_1 148
M2 fishfood_5 196
M2 leaves_2 218

15.07.2019
M2 oatmeal_19 135
```{r, message=FALSE, warning=FALSE}
remove <- c(148, 196, 184)

spider_treatment_M2 <- filter(spider_treatment_M2, date_endpoint != "2019-07-29" | spider_ID %notin% remove)

remove <- c(135)

spider_treatment_M2 <- filter(spider_treatment_M2, date_endpoint != "2019-07-15" | spider_ID %notin% remove)
```

No endpoints taken, because spiders were already dead

29.07.2019
M2 fishfood_1 198
M2 leaves_2 197
```{r, message=FALSE, warning=FALSE}
remove <- c(198, 197)

spider_treatment_M2 <- filter(spider_treatment_M2, date_endpoint != "2019-07-29" | spider_ID %notin% remove)
```

No endpoint taken

15.07.2019
M2 oatmeal_19 217
```{r, message=FALSE, warning=FALSE}
remove <- c(217)

spider_treatment_M2 <- filter(spider_treatment_M2, date_endpoint != "2019-07-15" | spider_ID %notin% remove)
```

## PUFA and dry weight data

* dry_pufa: Weight PUFA per dry weight of organism (microgramm PUFA per milligram dry mass)

Correct oatmealmeal_17
```{r, message=FALSE, warning=FALSE}
microcosm$experiment_ID <- gsub("oatmealmeal_17", "oatmeal_17", microcosm$experiment_ID)
```

Chose PUFA (= FA with more than one double bound)
```{r, message=FALSE, warning=FALSE}
only_pufa <- c("18:0", "18:1n-7", "18:1n-9c", "18:1n-9t", "20:0", "20:1n-9", "22:0", "22:1n-9", "24:0", "24:1n-9")

microcosm <- filter(microcosm, name_pufa %notin% only_pufa)
```

Calculate total PUFA amount and add to microcosm-table
```{r, message=FALSE, warning=FALSE}
microcosm_prop <- ddply(microcosm, c("Organism", "treatment", "experiment", "experiment_ID", "runinfo", "runname", "dry_weight_mg"), summarise, total_pufa=sum(dry_pufa))

microcosm <- merge(microcosm, microcosm_prop, by = c("Organism", "treatment", "experiment", "experiment_ID", "runinfo", "runname", "dry_weight_mg"))
```

Calculate proportion single PUFA of total PUFA
```{r, message=FALSE, warning=FALSE}
microcosm$prop_pufa <- with(microcosm, (dry_pufa/total_pufa))
```

Get date from runinfo: Separate column runinfo
```{r, message=FALSE, warning=FALSE}
microcosm <- separate(microcosm, runinfo, into = c("a", "b", "c", "d", "e"), sep = " ", remove = FALSE)
```

Check in which columns date
```{r, message=FALSE, warning=FALSE}
unique(microcosm$a) # no date
unique(microcosm$b) # only 06.05.2019
unique(microcosm$c) # several dates
unique(microcosm$d) # several dates
unique(microcosm$e) # only 31.05.2019
```

Get column date
```{r, message=FALSE, warning=FALSE}
microcosm$date <- with(microcosm, ifelse(!is.na(as.Date(microcosm$b,
                                                     "%d.%m.%Y")), b, NA))
microcosm$date <- with(microcosm, ifelse(!is.na(as.Date(microcosm$c,
                                                      "%d.%m.%Y")), c, date))
microcosm$date <- with(microcosm, ifelse(!is.na(as.Date(microcosm$d,
                                                      "%d.%m.%Y")), d, date))
microcosm$date <- with(microcosm, ifelse(!is.na(as.Date(microcosm$e,
                                                      "%d.%m.%Y")), e, date))

unique((filter(microcosm, is.na(date)))$Organism) # all NAs were spiders and this is correct, because there no date added
```

Remove columns a - e
```{r, message=FALSE, warning=FALSE}
microcosm <- select(microcosm, -c("a", "b", "c", "d", "e"))
```

Put in wide format for statistics
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- reshape2::dcast(microcosm, date + Organism + treatment + experiment + experiment_ID + runinfo + runinfo_original + runname + dry_weight_mg ~ name_pufa,
                       value.var = "prop_pufa", fill = 0)
```

Correct organism for PUFA-KF_25.01.2020 10_28_463
```{r, message=FALSE, warning=FALSE}
microcosm_stat$Organism[microcosm_stat$runname == "PUFA-KF_25.01.2020 10_28_463"]  <- "oatmeal"
```

Spiders:
* 176_leaves_18: PUFA-KF_08.01.2020 18_31_4311 and PUFA-KF_08.01.2020 19_27_2212. Keep second measurement, because first did not work 
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_08.01.2020 18_31_4311")
```

* 186_oatmeal_3: 4 measurements to test reaction time and volume sample and TMSH. Keep PUFA-KF_08.01.2020 17_34_5410, because this volumes and times used for other samples
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_08.01.2020 12_25_237")
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_08.01.2020 14_58_409")
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_08.01.2020 13_52_298")
```

Food sources:

* 31.05.2019 algae 2: Keep PUFA-KF_10.01.2020 21_34_4214, because better measurement
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_10.01.2020 16_00_528")
```

* 24.05.2019 algae 1: Keep PUFA-KF_30.05.2020 12_59_2929, because other measurement did not work properly
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_30.05.2020 09_16_2725")
```

* 06.05.2019 oatmeal: Keep PUFA-KF_10.01.2020 13_13_465, because better measurement
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_10.01.2020 18_47_4911")
```

* 04.06.2019 leaves 10: Keep PUFA-KF_10.01.2020 15_05_227, because better measurement
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_10.01.2020 20_39_0813")
```

* 24.05.2019 fishfood 2: Keep PUFA-KF_10.01.2020 14_09_396, because better measurement
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- filter(microcosm_stat, runname != "PUFA-KF_10.01.2020 19_43_2812")
```

Get variables for statistics
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- select(microcosm_stat, -c("runinfo", "runinfo_original", "runname"))
```

Get spider_ID_experiment_ID
```{r, message=FALSE, warning=FALSE}
microcosm_stat$spider_ID_experiment_ID <- with(microcosm_stat, ifelse(Organism == "spider" & treatment != "start" & experiment == "M2", experiment_ID, ""))
```

Get experiment_ID without spider_ID
```{r, message=FALSE, warning=FALSE}
microcosm_stat$experiment_ID <- gsub("^\\d\\d\\d[[:punct:]]", "", microcosm_stat$experiment_ID)
```

Merge with spider_treatment
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- left_join(microcosm_stat, spider_treatment, 
                  by = c("Organism", "treatment", "experiment", "experiment_ID"))
unique((filter(microcosm_stat, is.na(date_treatment_ended)))$treatment) # all NAs for spiders were start spiders and this is correct; all other NAs belong not so spiders so also correct
```

Add date for spiders
```{r, message=FALSE, warning=FALSE}
microcosm_stat$date <- with(microcosm_stat,
                           ifelse(Organism == "spider", date_treatment_ended, date))
```

Select columns
```{r, message=FALSE, warning=FALSE}
microcosm_stat <- select(microcosm_stat, c("date", "Organism", "treatment", "experiment", "experiment_ID", "dry_weight_mg", "18:2n-6c", "18:2n-6t", "18:3n-3", "18:3n-6", "20:2n-6", "20:3n-3", "20:3n-6", "20:4n-6", "20:5n-3", "22:2n-6", "22:6n-3", "spider_ID_experiment_ID"))
```

date in date format
```{r, message=FALSE, warning=FALSE}
microcosm_stat$date <- as.Date(microcosm_stat$date, "%d.%m.%Y")
names(microcosm_stat)[1] <- "date_endpoint"
```

## Remove spiders not used in experiment

Spiders with ID 101, 125, 128, 129, 141, 205 and 224 not used in experiment
```{r, message=FALSE, warning=FALSE}
remove <- c(101, 125, 128, 129, 141, 205, 224)

encapsulation_stat <- filter(encapsulation_stat, spider_ID %notin% remove)
growth_stat <- filter(growth_stat, spider_ID %notin% remove)
condition_stat <- filter(condition_stat, spider_ID %notin% remove)

microcosm_stat <- filter(microcosm_stat, experiment_ID != 141)
```

# Merge data
### Weekly endpoints

Weekly endpoints: growth_stat
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1 <- left_join(spider_treatment_M1, growth_stat, by = c("Organism", "date_endpoint", "experiment", "experiment_ID", "treatment", "spider_ID"))

spider_treatment_M2 <- left_join(spider_treatment_M2, growth_stat, by = c("Organism", "date_endpoint", "experiment", "experiment_ID", "treatment", "spider_ID"))
```

Weekly endpoints: condition_stat
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1 <- left_join(spider_treatment_M1, condition_stat, by = c("Organism", "date_endpoint", "experiment", "experiment_ID", "treatment", "spider_ID"))

spider_treatment_M2 <- left_join(spider_treatment_M2, condition_stat, by = c("Organism", "date_endpoint", "experiment", "experiment_ID", "treatment", "spider_ID"))
```

No condition pictures taken, because spider already broken:
2019-07-26	spider	M2	fishfood_19	fishfood	177: Only prosoma left
2019-07-26	spider	M2	fishfood_12	fishfood	166: Already dried

### Deadly endpoints

Deadly endpoints: encapsulation_stat
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1 <- left_join(spider_treatment_M1, encapsulation_stat, by = c("Organism", "date_endpoint", "experiment", "experiment_ID", "treatment", "spider_ID"))

spider_treatment_M2 <- left_join(spider_treatment_M2, encapsulation_stat, by = c("Organism", "date_endpoint", "experiment", "experiment_ID", "treatment", "spider_ID"))
```

Spider found dead, so no encapsulation assay:
M1 oatmeal_5 54
M1 oatmeal_5 123
M1 oatmeal_19 56
M1 fishfood_8 71
M1 fishfood_20 67
M1 fishfood_14 117
M1 fishfood_1 93
M1 algae_5 82
M1 algae_1 104

Spider not found, so no encapsulation assay:
M2 leaves_10 174

Deadly endpoints: microcosm_stat
```{r, message=FALSE, warning=FALSE}
spider_treatment_M1 <- left_join(spider_treatment_M1, microcosm_stat, by = c("Organism", "date_endpoint", "spider_ID_experiment_ID", "experiment", "experiment_ID", "treatment"))

spider_treatment_M2 <- left_join(spider_treatment_M2, microcosm_stat, by = c("Organism", "date_endpoint", "spider_ID_experiment_ID", "experiment", "experiment_ID", "treatment"))
```

# Deal with replacement spiders

Remove spiders, that died and were replaced:

* In M1 treatments with replacement: oatmeal_1, oatmeal_5, fishfood_7, algae_4, algae_19
```{r, message=FALSE, warning=FALSE}
remove <- c(27, 54, 84, 88, 105)
M1_stat <- filter(spider_treatment_M1, spider_ID %notin% remove)
```

* In M2 treatments with 2 replacements: oatmeal_2, fishfood_4. 
* Do not use replace_2 spider 182 from oatmeal_2, because many missing values and PUFA analysis was done with replace_1 spider. So remove replace_2 spider and original spider.

```{r, message=FALSE, warning=FALSE}
remove <- c(171, 182) 
M2_stat <- filter(spider_treatment_M2, spider_ID %notin% remove)
```

Use fishfood_4 replace_2, so remove replace_1 and original spider
```{r, message=FALSE, warning=FALSE}
remove <- c(175, 136)
M2_stat <- filter(M2_stat, spider_ID %notin% remove)
```

In M2 treatments with 1 replacement: oatmeal_19, fishfood_1, fishfood_2, fishfood_4, fishfood_12, fishfood_19, leaves_1, leaves_2, leaves_11, leaves_15, leaves_17, leaves_19, leaves_20.
```{r, message=FALSE, warning=FALSE}
remove <- c(135, 198, 170, 214, 199, 207, 197, 179, 183, 144, 210, 192) 
M2_stat <- filter(M2_stat, spider_ID %notin% remove)
```

# Prepare data for statistics
## Define time points 

M1: For replace spiders no values on 29.04.2019: Use 03.05.2019 as starting value
```{r, message=FALSE, warning=FALSE}
M1_stat$timepoint <- with(M1_stat, ifelse(date_endpoint == "2019-04-29", "t0",
                         ifelse(date_endpoint == "2019-05-08", "t2",
                         ifelse(date_endpoint == "2019-05-15", "t3",
                         ifelse(date_endpoint == "2019-05-22", "t4",
                         ifelse(date_endpoint == "2019-05-29", "t5",
                         ifelse(date_endpoint == "2019-06-05", "t6",
                         ifelse(date_endpoint == "2019-06-12", "t7",
                         ifelse(date_endpoint == "2019-05-03", "t0",
                                "no_regular")))))))))

```

Remove t0 = 29.04.2019 for replace spiders
```{r, message=FALSE, warning=FALSE}
remove <- c(123, 126, 127, 122, 124)
M1_stat <- filter(M1_stat, date_endpoint != "2019-04-29" | spider_ID %notin% remove)
```

Define time points for M2
```{r, message=FALSE, warning=FALSE}
M2_stat$timepoint <- with(M2_stat, ifelse(date_endpoint == "2019-07-08", "t0",
                         ifelse(date_endpoint == "2019-07-15", "t1",
                         ifelse(date_endpoint == "2019-07-22", "t2",
                         ifelse(date_endpoint == "2019-07-29", "t3",
                                "no_regular")))))
```

## Get one table for weekly and deadly endpoints

One table for weekly and one for deadly endpoints
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly <- filter(M1_stat, timepoint != "no_regular")
deadly <- c("t0", "t4", "t7")
M1_stat_deadly <- filter(M1_stat, timepoint %in% deadly)

M2_stat_weekly <- filter(M2_stat, timepoint != "no_regular")
M2_stat_deadly <- filter(M2_stat_weekly, timepoint != "t1")
```

Chose column for weekly and deadly endpoints
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly <- select(M1_stat_weekly, c("Organism", "treatment", "experiment",
                  "experiment_ID", "spider_ID", "weight_spider_g", 
                  "prosoma_width_cm", "opisthosoma_width_cm", 
                  "timepoint", "date_endpoint", "body_cond"))

M1_stat_deadly <- select(M1_stat_deadly, c("Organism", "treatment", "experiment",
                  "experiment_ID", "spider_ID", "timepoint", "dead_not_found_encapsulation","mean_R_e_ne",
                  "mean_area_encapsulation", "dry_weight_mg", 
                  "18:2n-6c", "18:2n-6t", "18:3n-3",
                  "18:3n-6", "20:2n-6", "20:3n-3", "20:3n-6",
                  "20:4n-6", "20:5n-3", "22:2n-6", "22:6n-3"))

M2_stat_weekly <- select(M2_stat_weekly, c("Organism", "treatment", "experiment",
                  "experiment_ID", "spider_ID", "weight_spider_g",  
                  "prosoma_width_cm", "opisthosoma_width_cm",
                  "timepoint", "date_endpoint", "body_cond"))

M2_stat_deadly <- select(M2_stat_deadly, c("Organism", "treatment", "experiment",
                  "experiment_ID", "spider_ID", "timepoint", "dead_not_found_encapsulation", "mean_R_e_ne",
                  "mean_area_encapsulation", "dry_weight_mg", 
                  "18:2n-6c", "18:2n-6t", "18:3n-3",
                  "18:3n-6", "20:2n-6", "20:3n-3", "20:3n-6",
                  "20:4n-6", "20:5n-3", "22:2n-6", "22:6n-3"))
```

## Deal with start spiders
### Deadly endpoints
Remove not start spiders from t0 deadly endpoints (only start spiders used there for deadly endpoints)
```{r, message=FALSE, warning=FALSE}
M1_stat_deadly <- filter(M1_stat_deadly, timepoint != "t0" | treatment == "start")

M2_stat_deadly <- filter(M2_stat_deadly, timepoint != "t0" | treatment == "start")
```

Remove spiders, from earlier timepoint, if they were used in later deadly endpoint
```{r, message=FALSE, warning=FALSE}
remove <- filter(M1_stat_deadly, timepoint == "t7")
remove <- unique(remove$spider_ID)
M1_stat_deadly <- filter(M1_stat_deadly, timepoint != "t4" | spider_ID %notin% remove)

remove <- filter(M2_stat_deadly, timepoint == "t3")
remove <- unique(remove$spider_ID)
M2_stat_deadly <- filter(M2_stat_deadly, timepoint != "t2" | spider_ID %notin% remove)
```

Assign start spiders randomly to treatment: M1
```{r, message=FALSE, warning=FALSE}
M1_start <- filter(M1_stat_deadly, treatment == "start")
treatment <- c(rep("algae", 7), rep("oatmeal", 7), rep("fishfood", 6))

set.seed(222)
M1_start$rnd_treatment <- sample(treatment)

M1_start_algae <- M1_start$spider_ID[M1_start$rnd_treatment == "algae"]
M1_start_oatmeal <- M1_start$spider_ID[M1_start$rnd_treatment == "oatmeal"]
M1_start_fishfood <- M1_start$spider_ID[M1_start$rnd_treatment == "fishfood"]

M1_stat_deadly$treatment <- with(M1_stat_deadly, 
                                 ifelse(spider_ID %in% M1_start_algae, "algae",
                                 ifelse(spider_ID %in% M1_start_oatmeal, "oatmeal",
                                 ifelse(spider_ID %in% M1_start_fishfood, "fishfood",
                                        treatment))))
```

Assign start spiders randomly to treatment: M2
```{r, message=FALSE, warning=FALSE}
M2_start <- filter(M2_stat_deadly, treatment == "start")
treatment <- c(rep("leaves", 5), rep("oatmeal", 4), rep("fishfood", 4))

set.seed(222)
M2_start$rnd_treatment <- sample(treatment)

M2_start_leaves <- M2_start$spider_ID[M2_start$rnd_treatment == "leaves"]
M2_start_oatmeal <- M2_start$spider_ID[M2_start$rnd_treatment == "oatmeal"]
M2_start_fishfood <- M2_start$spider_ID[M2_start$rnd_treatment == "fishfood"]

M2_stat_deadly$treatment <- with(M2_stat_deadly, 
                                 ifelse(spider_ID %in% M2_start_leaves, "leaves",
                                 ifelse(spider_ID %in% M2_start_oatmeal, "oatmeal",
                                 ifelse(spider_ID %in% M2_start_fishfood, "fishfood",
                                        treatment))))
```

Remove spiders, where no deadly endpoints taken, because spider was not found or already dead
```{r, message=FALSE, warning=FALSE}
M1_stat_deadly <- filter(M1_stat_deadly, dead_not_found_encapsulation == "e")

M2_stat_deadly <- filter(M2_stat_deadly, dead_not_found_encapsulation == "e")
```

### Weekly endpoints

Remove start spiders from weekly endpoints, because there for every spider variables measured for t0
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly <- filter(M1_stat_weekly, treatment != "start")

M2_stat_weekly <- filter(M2_stat_weekly, treatment != "start")
```

## Check NAs

Check NAs in M1_stat_deadly

* NA in encapsulation correct because monofilament lost: 22, 65, 44, 115, 24, 60 oatmeal 10, 90 oatmeal 20 
```{r, message=FALSE, warning=FALSE}
M1_NA_check <- filter(M1_stat_deadly, is.na(mean_R_e_ne))
filter(M1_NA_check, spider_ID %notin% c(22, 65, 44, 115, 24, 60, 90))
```

Check NAs in M2_stat_deadly

* NA in encapsulation correct because monofilament lost: 186 oatmeal 3, 131 fishfood 14
```{r, message=FALSE, warning=FALSE}
M2_NA_check <- filter(M2_stat_deadly, is.na(mean_R_e_ne))
filter(M2_NA_check, spider_ID %notin% c(186, 131))
```

Check NAs in M1_stat_weekly

* NA in opisthosoma_width_cm for algae_8 89 at t5 correctly, because opisthosoma was missing, so not use weight_spider_g from t5 for statistics
```{r, message=FALSE, warning=FALSE}
M1_NA_check <- filter(M1_stat_weekly, is.na(weight_spider_g))

M1_NA_check <- filter(M1_stat_weekly, is.na(prosoma_width_cm))

M1_NA_check <- filter(M1_stat_weekly, is.na(opisthosoma_width_cm))
```

Check NAs in M2_stat_weekly

* NAs fishfood_4 (two times) and leaves_2 correctly, no weight_spider_g and body_cond
```{r, message=FALSE, warning=FALSE}
M2_NA_check <- filter(M2_stat_weekly, is.na(weight_spider_g))

M2_NA_check <- filter(M2_stat_weekly, is.na(prosoma_width_cm))

M2_NA_check <- filter(M2_stat_weekly, is.na(opisthosoma_width_cm))
```

Remove fishfood_4 (two times) and leaves_2 
```{r, message=FALSE, warning=FALSE}
M2_stat_weekly <- filter(M2_stat_weekly, !is.na(weight_spider_g))
```

## Remove spiders with missing parts
Weight cannot be compared with other spiders and no photo for body condition could be taken 

Check spiders with missing opisthosoma
```{r, message=FALSE, warning=FALSE}
missing_opisthosoma <- filter(growth, comments %like% "opisthosoma")
missing_prosoma <- filter(growth, comments %like% "prosoma") # already removed earlier
```

algae_8 89 at t5, because opisthosoma was missing
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly <- filter(M1_stat_weekly, timepoint != "t5" | spider_ID %notin% c(89))
```

oatmeal_7 80 at t6, because opisthosoma was missing
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly <- filter(M1_stat_weekly, timepoint != "t6" | spider_ID %notin% c(80))
```

## Calculate growth rate

Check variability of fresh weight, thorax (= prosoma) as well as abdomen (=opisthosoma) width of spiders 

M1
```{r, message=FALSE, warning=FALSE}
M1_check_data <- reshape2::melt(M1_stat_weekly, id.vars = c("Organism", "treatment", "experiment", "experiment_ID", "spider_ID", "timepoint", "date_endpoint"))

M1_check_data <- filter(M1_check_data, variable != "body_cond")

ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="bottom", plot.title = element_text(size=20), legend.text = element_text(size=20), axis.text = element_text(size=15), axis.title = element_text(size=15)) +
geom_boxplot(data=M1_check_data[M1_check_data$variable == "weight_spider_g",], 
            aes(x=variable, y=value, col = treatment), show.legend = TRUE) +
  labs(title = "Fresh weight spider M1", 
       x="",
       y=expression(Weight ~"("~g~")")) 

ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="bottom", plot.title = element_text(size=20), legend.text = element_text(size=20), axis.text = element_text(size=15), axis.title = element_text(size=15)) +
geom_boxplot(data=M1_check_data[M1_check_data$variable != "weight_spider_g",], 
            aes(x=variable, y=value, col = treatment), show.legend = TRUE) +
  labs(title = "Width M1", 
       x="",
       y=expression(Width ~"("~cm~")")) 
```

M2
```{r, message=FALSE, warning=FALSE}
M2_check_data <- reshape2::melt(M2_stat_weekly, id.vars = c("Organism", "treatment", "experiment", "experiment_ID", "spider_ID", "timepoint", "date_endpoint"))

M2_check_data <- filter(M2_check_data, variable != "body_cond")

ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="bottom", plot.title = element_text(size=20), legend.text = element_text(size=20), axis.text = element_text(size=15), axis.title = element_text(size=15)) +
geom_boxplot(data=M2_check_data[M2_check_data$variable == "weight_spider_g",], 
            aes(x=variable, y=value, col = treatment), show.legend = TRUE) +
  labs(title = "Fresh weight spider M2", 
       x="",
       y=expression(Weight ~"("~g~")")) 

ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="bottom", plot.title = element_text(size=20), legend.text = element_text(size=20), axis.text = element_text(size=15), axis.title = element_text(size=15)) +
geom_boxplot(data=M2_check_data[M2_check_data$variable != "weight_spider_g",], 
            aes(x=variable, y=value, col = treatment), show.legend = TRUE) +
  labs(title = "Width M2", 
       x="",
       y=expression(Width ~"("~cm~")")) 
```

In both experiments abdomen slightly more variability than thorax

Prepare data

time point in days
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly$days <- with(M1_stat_weekly, ifelse(timepoint == "t0", 0,
                                            ifelse(timepoint == "t2", 9,
                                            ifelse(timepoint == "t3", 16,
                                            ifelse(timepoint == "t4", 23,
                                            ifelse(timepoint == "t5", 30,
                                            ifelse(timepoint == "t6", 37, 44)))))))

M2_stat_weekly$days <- with(M2_stat_weekly, ifelse(timepoint == "t0", 0,
                                            ifelse(timepoint == "t1", 7,
                                            ifelse(timepoint == "t2", 14, 21))))
```

time point in weeks
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly$weeks <- with(M1_stat_weekly, days/7)

M2_stat_weekly$weeks <- with(M2_stat_weekly, days/7)
```

Get fresh weight and thorax width at t0 for every spider
```{r, message=FALSE, warning=FALSE}
M1_t0 <- filter(M1_stat_weekly, timepoint == "t0")
M1_t0 <- select(M1_t0, c("Organism", "treatment", "experiment", "experiment_ID",
                         "spider_ID", "weight_spider_g", "prosoma_width_cm"))
names(M1_t0)[6] <- "weight_t0"
names(M1_t0)[7] <- "thorax_t0"

M1_stat_weekly <- merge(M1_stat_weekly, M1_t0, by = c("Organism", "treatment", "experiment", "experiment_ID", "spider_ID"))

M2_t0 <- filter(M2_stat_weekly, timepoint == "t0")
M2_t0 <- select(M2_t0, c("Organism", "treatment", "experiment", "experiment_ID",
                         "spider_ID", "weight_spider_g", "prosoma_width_cm"))
names(M2_t0)[6] <- "weight_t0"
names(M2_t0)[7] <- "thorax_t0"

M2_stat_weekly <- merge(M2_stat_weekly, M2_t0, by = c("Organism", "treatment", "experiment", "experiment_ID", "spider_ID"))
```



$$ growth\ rate = {{ln (fresh\ weight_{t_x}) - ln (fresh\ weight_{t_0})}  \over {t_x}} $$
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly$growth_rate <- with(M1_stat_weekly,
                                    (log(weight_spider_g)-log(weight_t0))/weeks)

M2_stat_weekly$growth_rate <- with(M2_stat_weekly,
                                    (log(weight_spider_g)-log(weight_t0))/weeks)
```

## Chose data for statistical analysis
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly <- select(M1_stat_weekly, c("Organism", "treatment", "experiment", "experiment_ID", "spider_ID", "weight_spider_g", "timepoint", "date_endpoint", "body_cond", "growth_rate"))

M2_stat_weekly <- select(M2_stat_weekly, c("Organism", "treatment", "experiment", "experiment_ID", "spider_ID", "weight_spider_g", "timepoint", "date_endpoint", "body_cond", "growth_rate"))
```

# NMDS
## Preparation

Use microcosm_stat, because we need also PUFA data for chiros and basic food source. On M1_stat_deadly and M2_stat_deadly only spiders.
```{r, message=FALSE, warning=FALSE}
M1_microcosm_stat <- filter(microcosm_stat, experiment == "M1")
M2_microcosm_stat <- filter(microcosm_stat, experiment == "M2")
```

Add time point M1
```{r, message=FALSE, warning=FALSE}
M1_microcosm_stat$timepoint <- with(M1_microcosm_stat, 
       ifelse(date_endpoint < "2019-05-03", "t0",
       ifelse("2019-05-03" <= date_endpoint & date_endpoint <= "2019-05-22", "t4",
               "t7")))
```

Add time point M2
For leaves date is time, when leaves were taken out of stream. Leaves used in experiment at other date/time point:

* 2019-06-04 and 2019-06-18: t2 
* 2019-07-02 and 2019-07-16: t3

```{r, message=FALSE, warning=FALSE}
M2_microcosm_stat$timepoint <- with(M2_microcosm_stat, 
       ifelse(date_endpoint == "2019-07-08", "t0",
       ifelse(date_endpoint == "2019-07-22", "t2",
       ifelse(date_endpoint == "2019-06-04", "t2", 
       ifelse(date_endpoint == "2019-06-18", "t2",
       ifelse(date_endpoint == "2019-07-12", "t2",
       ifelse(date_endpoint == "2019-07-19", "t2",      
       ifelse(date_endpoint == "2019-07-02", "t3",
       ifelse(date_endpoint == "2019-07-16", "t3"        
               ,"t3")))))))))  
```

Assign start spiders randomly to treatment: M1
```{r, message=FALSE, warning=FALSE}
M1_microcosm_stat$treatment <- with(M1_microcosm_stat, 
                                 ifelse(experiment_ID %in% M1_start_algae, "algae",
                                 ifelse(experiment_ID %in% M1_start_oatmeal, "oatmeal",
                                 ifelse(experiment_ID %in% M1_start_fishfood, "fishfood",
                                        treatment))))
```

Assign start spiders randomly to treatment: M2
```{r, message=FALSE, warning=FALSE}
M2_microcosm_stat$treatment <- with(M2_microcosm_stat, 
                                 ifelse(experiment_ID %in% M2_start_leaves, "leaves",
                                 ifelse(experiment_ID %in% M2_start_oatmeal, "oatmeal",
                                 ifelse(experiment_ID %in% M2_start_fishfood, "fishfood",
                                        treatment))))
```

Count number of samples per treatment, trophic level and endpoint
```{r, message=FALSE, warning=FALSE}
M1_count <- count(M1_microcosm_stat, c("Organism", "treatment", "timepoint"))
M1_count$timepoint <- with(M1_count, ifelse(timepoint == "t0", 0,
                                     ifelse(timepoint == "t4", 23, 44)))
M1_count$Organism <- gsub("chiro", "chironomid", M1_count$Organism)
M1_count$Organism <- with(M1_count, 
                     ifelse(Organism %in% 
                              c("algae", "fishfood", "oatmeal"), 
                            "basic food source",Organism))
M1_count$treatment <- gsub("fishfood", "fish food", M1_count$treatment)
M1_count$experiment <- "1"

M1_count <- arrange(M1_count, treatment, timepoint, Organism)

names(M1_count)[1] <- "trophic level"
names(M1_count)[3] <- "time point (days)"
names(M1_count)[4] <- "number samples"

M1_count <- select(M1_count, c("experiment", "treatment", 
                   "time point (days)", "trophic level", "number samples"))


M2_count <- count(M2_microcosm_stat, c("Organism", "treatment", "timepoint")) 
M2_count$timepoint <- with(M2_count, ifelse(timepoint == "t0", 0,
                                     ifelse(timepoint == "t2", 14, 21)))
M2_count$Organism <- gsub("chiro", "chironomid", M2_count$Organism)
M2_count$Organism <- with(M2_count, 
                     ifelse(Organism %in% 
                              c("leaves", "fishfood", "oatmeal"), 
                            "basic food source",Organism))
M2_count$treatment <- gsub("fishfood", "fish food", M2_count$treatment)
M2_count$experiment <- "2"

M2_count <- arrange(M2_count, treatment, timepoint, Organism)

names(M2_count)[1] <- "trophic level"
names(M2_count)[3] <- "time point (days)"
names(M2_count)[4] <- "number samples"

M2_count <- select(M2_count, c("experiment", "treatment", 
                   "time point (days)", "trophic level", "number samples"))


count <- rbind(M1_count, M2_count)

#write.csv(count, "no_samples_count.csv", row.names = FALSE)
```

Get matrix for pufa and treatment information
```{r, message=FALSE, warning=FALSE}
M1_treatment <- select(M1_microcosm_stat, c("Organism", "treatment", "timepoint"))

M1_pufa <- select(M1_microcosm_stat, -c("Organism", "treatment", "timepoint", "date_endpoint", "experiment", "experiment_ID", "dry_weight_mg", "spider_ID_experiment_ID"))
M1_pufa_ma <- as.matrix(M1_pufa)

M2_treatment <- select(M2_microcosm_stat, c("Organism", "treatment", "timepoint"))

M2_pufa <- select(M2_microcosm_stat, -c("Organism", "treatment", "timepoint", "date_endpoint", "experiment", "experiment_ID", "dry_weight_mg", "spider_ID_experiment_ID"))
M2_pufa_ma <- as.matrix(M2_pufa)
```

Function to calculate relationship between dimensions and stress
```{r, message=FALSE, warning=FALSE}
scree_values <- function(x, max, dist_meas) {
  xx <- as.matrix(x)
  scree.points <- NULL
  scree.temp <- 1
  for (i in 1:max)
  {
    sol.mds <- NULL
    sol.mds <- replicate(
      10, metaMDS(xx, k = i, trymax = 30, distance = dist_meas)$stress,
      simplify = TRUE
    )
    scree.points <- append(scree.points, sol.mds)
  }
  return(scree.points)
}
```

## M1

Get and plot relationship between dimensions and stress
* red line indicates threshold for good Stress1 value
* Choose 2 dimensions

```{r, message=FALSE, warning=FALSE}
#M1_scree.res <- scree_values(M1_pufa_ma, max = 10, dist_meas = "euclidean")
#saveRDS(M1_scree.res, file = "M1_scree.res.rds")
M1_scree.res <- readRDS(file = "/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/PUFA_data/M1_scree.res.rds")

par(cex = 1.5, las = 1)
plot_vec <- as.vector(sapply(1:(length(M1_scree.res) / 10), function(x) rep(x, 10), simplify = "vector"))
# prerequisite for plotting replicated runs of nmds
plot(plot_vec, M1_scree.res, ylab = "Stress1", xlab = "dimensions")
lines(1:(length(M1_scree.res) / 10), as.vector((by(M1_scree.res, plot_vec, mean))))
abline(h = 0.10, col = "red")
```

### Run NMDS 
with Euclidean distance
```{r, message=FALSE, warning=FALSE}
#M1_nmds <- metaMDS(M1_pufa_ma, k = 2, distance = "euclidean", autotransform = FALSE)
#saveRDS(M1_nmds, file = "M1_nmds.rds")
M1_nmds <- readRDS(file = "/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/PUFA_data/M1_nmds.rds")
```

Shepard diagram
* Non-metric fit is based on stress and calculated as R^2 = sqrt(1- S^2)
* Linear R^2 = explained variation of regression of observed distances against ordination distances
* Looks okay
```{r, message=FALSE, warning=FALSE}
stressplot(M1_nmds)
```

### Plot NMDS
#### Prepare data for plotting
Get data of NMDS
```{r, message=FALSE, warning=FALSE}
M1_pufa_scores <- as.data.frame(M1_nmds$species)
M1_pufa_scores$pufa <- rownames(M1_pufa_scores)

M1_sample_scores <-as.data.frame(M1_nmds$points)
M1_sample_scores <- cbind(M1_sample_scores, M1_treatment)
```

Rename some variables
```{r, message=FALSE, warning=FALSE}
M1_sample_scores$treatment <- gsub("fishfood", "fish food", M1_sample_scores$treatment)
M1_sample_scores$Organism <- gsub("fishfood", "fish food", M1_sample_scores$Organism)
M1_sample_scores$Organism <- gsub("chiro", "chironomids", M1_sample_scores$Organism)
M1_sample_scores$timepoint <- gsub("t0", "0 days", M1_sample_scores$timepoint)
M1_sample_scores$timepoint <- gsub("t4", "23 days", M1_sample_scores$timepoint)
M1_sample_scores$timepoint <- gsub("t7", "44 days", M1_sample_scores$timepoint)
```

Get single spider scores and mean and sd of chiro and food source scores
```{r, message=FALSE, warning=FALSE}
M1_spider_scores <- filter(M1_sample_scores, Organism == "spider")

M1_chiro_food <- filter(M1_sample_scores, Organism != "spider")
M1_chiro_food_mean <- ddply(M1_chiro_food, c("Organism", "treatment", "timepoint"),
                              summarise, MDS1 = mean(MDS1), MDS2 = mean(MDS2))

M1_chiro_food_sd <- filter(M1_sample_scores, Organism != "spider")
M1_chiro_food_sd <- ddply(M1_chiro_food_sd, c("Organism", "treatment", "timepoint"),
                              summarise, sd_MDS1 = sd(MDS1),sd_MDS2 = sd(MDS2))

M1_chiro_food_sd <- left_join(M1_chiro_food_mean, M1_chiro_food_sd, 
                           by = c("Organism", "treatment", "timepoint"))

M1_spider_chiro_food_scores <- rbind(M1_spider_scores, M1_chiro_food_mean)
```

Add column with trophic level
```{r, message=FALSE, warning=FALSE}
M1_spider_chiro_food_scores$trophic_level <- with(M1_spider_chiro_food_scores, ifelse(Organism == "spider", "spider", ifelse(Organism == "chironomids", "chironomids", "basic food source")))
```

Reorder factors
```{r, message=FALSE, warning=FALSE}
M1_spider_chiro_food_scores$treatment <- factor(M1_spider_chiro_food_scores$treatment, levels = c("algae", "oatmeal", "fish food"))

M1_chiro_food_sd$treatment <- factor(M1_chiro_food_sd$treatment, levels = c("algae", "oatmeal", "fish food"))

M1_spider_chiro_food_scores$Organism <- factor(M1_spider_chiro_food_scores$Organism, levels = c("spider", "chironomids", "algae", "oatmeal", "fish food"))

M1_chiro_food_sd$Organism <- factor(M1_chiro_food_sd$Organism, levels = c("spider", "chironomids", "algae", "oatmeal", "fish food"))

M1_spider_chiro_food_scores$trophic_level <- factor(M1_spider_chiro_food_scores$trophic_level, levels = c("spider", "chironomids", "basic food source"))
```

Rename PUFAs and chose ARA and PUFA contributing for 70 % differences between treatments (see SIMPER) for plotting (otherwise too unclear)
```{r, message=FALSE, warning=FALSE}
M1_pufa_scores_all <- M1_pufa_scores

M1_pufa_scores$pufa <- gsub("22:6n-3", "DHA", M1_pufa_scores$pufa) # docosahexaenoic acid
  
M1_pufa_scores$pufa <- gsub("20:5n-3", "EPA", M1_pufa_scores$pufa) # eicosapentaenoic acid
M1_pufa_scores$pufa <- gsub("20:4n-6", "ARA", M1_pufa_scores$pufa) # arachidonic acid

M1_pufa_scores$pufa <- gsub("18:3n-3", "ALA", M1_pufa_scores$pufa) # alpha-linolenic acid
M1_pufa_scores$pufa <- gsub("18:3n-6", "GLA", M1_pufa_scores$pufa) # gamma-Linolenic acid

M1_pufa_scores$pufa <- gsub("18:2n-6c", "LIN", M1_pufa_scores$pufa) # Linoleic acid
M1_pufa_scores$pufa <- gsub("18:2n-6t", "LLA", M1_pufa_scores$pufa) # Linolelaidic acid

M1_pufa_scores <- filter(M1_pufa_scores, pufa %in% c("ALA", "LIN", "EPA", "ARA", "GLA", "DHA",  "LLA" ))
```

#### Plot
##### Only PUFA contributing to differences between treatments
Results from SIMPER 
```{r, message=FALSE, warning=FALSE}
M1_nmds_plot <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="right", 
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text = element_text(size=15), axis.title = element_text(size=15),
      strip.text = element_text(size=15)) +
  labs(title = "Experiment 1", 
       x=expression(NMDS1),
       y=expression(NMDS2),
       shape ="trophic level") +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3")) +
  scale_shape_manual(values=c(21,22,24)) +
  facet_grid(~ timepoint) +
  geom_point(data = M1_spider_chiro_food_scores, 
             aes(x=MDS1, y=MDS2, col = treatment,
                 shape = trophic_level, fill = treatment),
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5) +
  geom_errorbar(data = M1_chiro_food_sd, 
                aes(x = MDS1, y = MDS2,
                    xmin=MDS1-sd_MDS1, xmax=MDS1+sd_MDS1,
                    col = treatment), show.legend = FALSE,
                size = 1, alpha = 0.5) +
  geom_errorbar(data = M1_chiro_food_sd, 
                aes(x = MDS1, y = MDS2,
                    ymin=MDS2-sd_MDS2, ymax=MDS2+sd_MDS2,
                    col = treatment), show.legend = FALSE,
                size = 1, alpha = 0.5) +
   geom_text_repel(data = M1_pufa_scores, aes(x = MDS1, y = MDS2, label = pufa),
                  size = 3, max.overlaps = 16, fontface = "bold", alpha = 0.5)

M1_nmds_plot
```

##### All PUFA
```{r, message=FALSE, warning=FALSE}
M1_nmds_plot_all <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="right", 
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text = element_text(size=15), axis.title = element_text(size=15),
      strip.text = element_text(size=15)) +
  labs(title = "Experiment 1", 
       x=expression(NMDS1),
       y=expression(NMDS2),
       shape ="trophic level") +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3")) +
  scale_shape_manual(values=c(21,22,24)) +
  facet_grid(~ timepoint) +
  geom_point(data = M1_spider_chiro_food_scores, 
             aes(x=MDS1, y=MDS2, col = treatment,
                 shape = trophic_level, fill = treatment),
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5) +
  geom_errorbar(data = M1_chiro_food_sd, 
                aes(x = MDS1, y = MDS2,
                    xmin=MDS1-sd_MDS1, xmax=MDS1+sd_MDS1,
                    col = treatment), show.legend = FALSE,
                size = 1, alpha = 0.5) +
  geom_errorbar(data = M1_chiro_food_sd, 
                aes(x = MDS1, y = MDS2,
                    ymin=MDS2-sd_MDS2, ymax=MDS2+sd_MDS2,
                    col = treatment), show.legend = FALSE,
                size = 1, alpha = 0.5) +
 geom_text_repel(data = M1_pufa_scores_all, aes(x = MDS1, y = MDS2, label = pufa),
                  size = 3, max.overlaps = 16, fontface = "bold", alpha = 0.5)

M1_nmds_plot_all
```


## M2

Get and plot relationship between dimensions and stress
* red line indicates threshold for good Stress1 value
* Choose 2 dimensions

```{r, message=FALSE, warning=FALSE}
#M2_scree.res <- scree_values(M2_pufa_ma, max = 10, dist_meas = "euclidean")
#saveRDS(M2_scree.res, file = "M2_scree.res.rds")
M2_scree.res <- readRDS(file = "/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/PUFA_data/M2_scree.res.rds")

par(cex = 1.5, las = 1)
plot_vec <- as.vector(sapply(1:(length(M2_scree.res) / 10), function(x) rep(x, 10), simplify = "vector"))
# prerequisite for plotting replicated runs of nmds
plot(plot_vec, M2_scree.res, ylab = "Stress1", xlab = "dimensions")
lines(1:(length(M2_scree.res) / 10), as.vector((by(M2_scree.res, plot_vec, mean))))
abline(h = 0.10, col = "red")
```

### Run NMDS
with Euclidean distance
```{r, message=FALSE, warning=FALSE}
#M2_nmds <- metaMDS(M2_pufa_ma, k = 2, distance = "euclidean", autotransform = FALSE)
#saveRDS(M2_nmds, file = "M2_nmds.rds")
M2_nmds <- readRDS(file = "/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Submission/Ecology_and_Evolution_submission/2_second_submission/ohler_microcosm/statistics/PUFA_data/M2_nmds.rds")
```

Shepard diagram
* Non-metric fit is based on stress and calculated as R^2 = sqrt(1- S^2)
* Linear R^2 = explained variation of regression of observed distances against ordination distances
* Looks okay
```{r, message=FALSE, warning=FALSE}
stressplot(M2_nmds)
```

### Plot NMDS
#### Prepare data for plotting
```{r, message=FALSE, warning=FALSE}
M2_pufa_scores <- as.data.frame(M2_nmds$species)
M2_pufa_scores$pufa <- rownames(M2_pufa_scores)

M2_sample_scores <-as.data.frame(M2_nmds$points)
M2_sample_scores <- cbind(M2_sample_scores, M2_treatment)
```

Rename some variables
```{r, message=FALSE, warning=FALSE}
M2_sample_scores$treatment <- gsub("fishfood", "fish food", M2_sample_scores$treatment)
M2_sample_scores$Organism <- gsub("fishfood", "fish food", M2_sample_scores$Organism)
M2_sample_scores$Organism <- gsub("chiro", "chironomids", M2_sample_scores$Organism)
M2_sample_scores$timepoint <- gsub("t0", "0 days", M2_sample_scores$timepoint)
M2_sample_scores$timepoint <- gsub("t2", "14 days", M2_sample_scores$timepoint)
M2_sample_scores$timepoint <- gsub("t3", "21 days", M2_sample_scores$timepoint)
```

Get single spider scores and mean and sd of chiro and food source scores
```{r, message=FALSE, warning=FALSE}
M2_spider_scores <- filter(M2_sample_scores, Organism == "spider")

M2_chiro_food <- filter(M2_sample_scores, Organism != "spider")
M2_chiro_food_mean <- ddply(M2_chiro_food, c("Organism", "treatment", "timepoint"),
                              summarise, MDS1 = mean(MDS1), MDS2 = mean(MDS2))

M2_chiro_food_sd <- filter(M2_sample_scores, Organism != "spider")
M2_chiro_food_sd <- ddply(M2_chiro_food_sd, c("Organism", "treatment", "timepoint"),
                              summarise, sd_MDS1 = sd(MDS1),sd_MDS2 = sd(MDS2))

M2_chiro_food_sd <- left_join(M2_chiro_food_mean, M2_chiro_food_sd, 
                           by = c("Organism", "treatment", "timepoint"))

M2_spider_chiro_food_scores <- rbind(M2_spider_scores, M2_chiro_food_mean)
```

Add column with trophic level
```{r, message=FALSE, warning=FALSE}
M2_spider_chiro_food_scores$trophic_level <- with(M2_spider_chiro_food_scores, ifelse(Organism == "spider", "spider", ifelse(Organism == "chironomids", "chironomids", "basic food source")))
```

Reorder factors
```{r, message=FALSE, warning=FALSE}
M2_spider_chiro_food_scores$treatment <- factor(M2_spider_chiro_food_scores$treatment, levels = c("leaves", "oatmeal", "fish food"))

M2_chiro_food_sd$treatment <- factor(M2_chiro_food_sd$treatment, levels = c("leaves", "oatmeal", "fish food"))

M2_spider_chiro_food_scores$Organism <- factor(M2_spider_chiro_food_scores$Organism, levels = c("spider", "chironomids", "leaves", "oatmeal", "fish food"))

M2_chiro_food_sd$Organism <- factor(M2_chiro_food_sd$Organism, levels = c("spider", "chironomids", "leaves", "oatmeal", "fish food"))

M2_spider_chiro_food_scores$trophic_level <- factor(M2_spider_chiro_food_scores$trophic_level, levels = c("spider", "chironomids", "basic food source"))
```

Rename PUFAs and chose ARA and PUFA contributing for 70 % differences between treatments (see SIMPER) for plotting (otherwise too unclear)
```{r, message=FALSE, warning=FALSE}
M2_pufa_scores_all <- M2_pufa_scores

M2_pufa_scores$pufa <- gsub("22:6n-3", "DHA", M2_pufa_scores$pufa) # docosahexaenoic acid
  
M2_pufa_scores$pufa <- gsub("20:5n-3", "EPA", M2_pufa_scores$pufa) # eicosapentaenoic acid
M2_pufa_scores$pufa <- gsub("20:4n-6", "ARA", M2_pufa_scores$pufa) # arachidonic acid

M2_pufa_scores$pufa <- gsub("18:3n-3", "ALA", M2_pufa_scores$pufa) # alpha-linolenic acid
M2_pufa_scores$pufa <- gsub("18:3n-6", "GLA", M2_pufa_scores$pufa) # gamma-Linolenic acid

M2_pufa_scores$pufa <- gsub("18:2n-6c", "LIN", M2_pufa_scores$pufa) # Linoleic acid
M2_pufa_scores$pufa <- gsub("18:2n-6t", "LLA", M2_pufa_scores$pufa) # Linolelaidic acid

M2_pufa_scores <- filter(M2_pufa_scores, pufa %in% c("ALA", "LIN", "EPA", "ARA", "GLA", "DHA",  "LLA" ))
```

#### Plot
##### Only Only PUFA contributing to differences between treatments
```{r, message=FALSE, warning=FALSE}
M2_nmds_plot <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="right", 
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text = element_text(size=15), axis.title = element_text(size=15),
      strip.text = element_text(size=15)) +
  labs(title = "Experiment 2", 
       x = expression(NMDS1),
       y = expression(NMDS2),
       shape = "trophic level") +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3")) +
  scale_shape_manual(values=c(21, 22, 24)) +
  facet_grid(~ timepoint) +
  geom_point(data = M2_spider_chiro_food_scores, 
             aes(x=MDS1, y=MDS2, col = treatment,
                 shape = trophic_level, fill = treatment),
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5) +
  geom_errorbar(data = M2_chiro_food_sd, 
                aes(x = MDS1, y = MDS2, 
                    xmin=MDS1-sd_MDS1, xmax=MDS1+sd_MDS1,
                    col = treatment), show.legend = FALSE,
                size = 1, alpha = 0.5) +
  geom_errorbar(data = M2_chiro_food_sd, 
                aes(x = MDS1, y = MDS2,
                    ymin=MDS2-sd_MDS2, ymax=MDS2+sd_MDS2,
                    col = treatment), show.legend = FALSE,
                size = 1, alpha = 0.5) +
 geom_text_repel(data = M2_pufa_scores, aes(x = MDS1, y = MDS2, label = pufa),
                  size = 3, max.overlaps = 16, fontface = "bold", alpha = 0.5)

M2_nmds_plot
```

##### All PUFA
```{r, message=FALSE, warning=FALSE}
M2_nmds_plot_all <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="right", 
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text = element_text(size=15), axis.title = element_text(size=15),
      strip.text = element_text(size=15)) +
  labs(title = "Experiment 2", 
       x = expression(NMDS1),
       y = expression(NMDS2),
       shape = "trophic level") +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3")) +
  scale_shape_manual(values=c(21, 22, 24)) +
  facet_grid(~ timepoint) +
  geom_point(data = M2_spider_chiro_food_scores, 
             aes(x=MDS1, y=MDS2, col = treatment,
                 shape = trophic_level, fill = treatment),
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5) +
  geom_errorbar(data = M2_chiro_food_sd, 
                aes(x = MDS1, y = MDS2, 
                    xmin=MDS1-sd_MDS1, xmax=MDS1+sd_MDS1,
                    col = treatment), show.legend = FALSE,
                size = 1, alpha = 0.5) +
  geom_errorbar(data = M2_chiro_food_sd, 
                aes(x = MDS1, y = MDS2,
                    ymin=MDS2-sd_MDS2, ymax=MDS2+sd_MDS2,
                    col = treatment), show.legend = FALSE,
                size = 1, alpha = 0.5) +
 geom_text_repel(data = M2_pufa_scores_all, aes(x = MDS1, y = MDS2, label = pufa),
                  size = 3, max.overlaps = 16, fontface = "bold", alpha = 0.5)

M2_nmds_plot_all
```

## Save NMDS plots
```{r, message=FALSE, warning=FALSE}
NMDS_plot <- ggarrange(M1_nmds_plot, M2_nmds_plot,
                  labels = c("A", "B"),
                  ncol = 1, nrow = 2)

#png(filename="NMDS_plot.png", width=15, height=10, units="in", res=300)
NMDS_plot
#dev.off()

NMDS_plot_all <- ggarrange(M1_nmds_plot_all, M2_nmds_plot_all,
                  labels = c("A", "B"),
                  ncol = 1, nrow = 2)

#png(filename="NMDS_plot_all.png", width=15, height=10, units="in", res=300)
NMDS_plot_all
#dev.off()
```

# ANOSIM
## M1
### Spider within same time point: compare treatments
```{r, message=FALSE, warning=FALSE}
spider_t4 <- which(M1_treatment$timepoint == "t4" & M1_treatment$Organism == "spider")
set.seed(222)
M1_anosim_spider_t4 <- anosim(M1_pufa_ma[spider_t4,], M1_treatment[spider_t4,]$treatment, distance = "euclidean")
summary(M1_anosim_spider_t4)

spider_t7 <- which(M1_treatment$timepoint == "t7" & M1_treatment$Organism == "spider")
set.seed(222)
M1_anosim_spider_t7 <- anosim(M1_pufa_ma[spider_t7,], M1_treatment[spider_t7,]$treatment, distance = "euclidean")
summary(M1_anosim_spider_t7)
```

### Spider within same treatment and start spiders: compare time points
```{r, message=FALSE, warning=FALSE}
M1_spider_algae_time <- which(M1_treatment$Organism == "spider" & M1_treatment$treatment %in% c("algae"))

set.seed(222)
M1_anosim_spider_algae_time <- anosim(M1_pufa_ma[M1_spider_algae_time,], M1_treatment[M1_spider_algae_time,]$timepoint, distance = "euclidean")
summary(M1_anosim_spider_algae_time)

M1_spider_oatmeal_time <- which(M1_treatment$Organism == "spider" & M1_treatment$treatment %in% c("oatmeal"))

set.seed(222)
M1_anosim_spider_oatmeal_time <- anosim(M1_pufa_ma[M1_spider_oatmeal_time,], M1_treatment[M1_spider_oatmeal_time,]$timepoint, distance = "euclidean")
summary(M1_anosim_spider_oatmeal_time)

M1_spider_fishfood_time <- which(M1_treatment$Organism == "spider" & M1_treatment$treatment %in% c("fishfood"))

set.seed(222)
M1_anosim_spider_fishfood_time <- anosim(M1_pufa_ma[M1_spider_fishfood_time,], M1_treatment[M1_spider_fishfood_time,]$timepoint, distance = "euclidean")
summary(M1_anosim_spider_fishfood_time)
```

### Chiro within same time point: compare treatments
```{r, message=FALSE, warning=FALSE}
chiro_t4 <- which(M1_treatment$timepoint == "t4" & M1_treatment$Organism == "chiro")
set.seed(222)
M1_anosim_chiro_t4 <- anosim(M1_pufa_ma[chiro_t4,], M1_treatment[chiro_t4,]$treatment, distance = "euclidean")
summary(M1_anosim_chiro_t4)

chiro_t7 <- which(M1_treatment$timepoint == "t7" & M1_treatment$Organism == "chiro")
set.seed(222)
M1_anosim_chiro_t7 <- anosim(M1_pufa_ma[chiro_t7,], M1_treatment[chiro_t7,]$treatment, distance = "euclidean")
summary(M1_anosim_chiro_t7)
```

### Food source within same time point: compare treatments
```{r, message=FALSE, warning=FALSE}
food_t4 <- which(M1_treatment$timepoint == "t4" & M1_treatment$Organism %in% c("algae", "oatmeal", "fishfood"))

set.seed(222)
M1_anosim_food_t4 <- anosim(M1_pufa_ma[food_t4,], M1_treatment[food_t4,]$treatment, distance = "euclidean")
summary(M1_anosim_food_t4)

food_t7 <- which(M1_treatment$timepoint == "t7" & M1_treatment$Organism %in% c("algae", "oatmeal", "fishfood"))

set.seed(222)
M1_anosim_food_t7 <- anosim(M1_pufa_ma[food_t7,], M1_treatment[food_t7,]$treatment, distance = "euclidean")
summary(M1_anosim_food_t7)
```

### Extract R and p-values
```{r, message=FALSE, warning=FALSE}
M1_anosim_name <- c("M1_anosim_spider_t4", "M1_anosim_spider_t7",
                    "M1_anosim_spider_algae_time", "M1_anosim_spider_oatmeal_time",
                    "M1_anosim_spider_fishfood_time", "M1_anosim_chiro_t4",
                    "M1_anosim_chiro_t7", "M1_anosim_food_t4", "M1_anosim_food_t7")

M1_anosim_p <- rbind(M1_anosim_spider_t4$signif, M1_anosim_spider_t7$signif,
                 M1_anosim_spider_algae_time$signif, M1_anosim_spider_oatmeal_time$signif,
                 M1_anosim_spider_fishfood_time$signif, M1_anosim_chiro_t4$signif,
                 M1_anosim_chiro_t7$signif, M1_anosim_food_t4$signif,
                 M1_anosim_food_t7$signif)

M1_anosim_R <- rbind(M1_anosim_spider_t4$statistic, M1_anosim_spider_t7$statistic,
                 M1_anosim_spider_algae_time$statistic,
                 M1_anosim_spider_oatmeal_time$statistic,
                 M1_anosim_spider_fishfood_time$statistic, M1_anosim_chiro_t4$statistic,
                 M1_anosim_chiro_t7$statistic, M1_anosim_food_t4$statistic,
                 M1_anosim_food_t7$statistic)

M1_anosim <- as.data.frame(cbind(M1_anosim_name, M1_anosim_p, M1_anosim_R))

names(M1_anosim)[2] <- "p_value"
names(M1_anosim)[3] <- "R"
```

### Adjust p-values and save output
```{r, message=FALSE, warning=FALSE}
M1_anosim$p_value_adjusted <- p.adjust(M1_anosim$p_value, method = "hochberg")

#write.csv(M1_anosim, "M1_anosim.csv", row.names = FALSE)
```

## M2
### Spider within same time point: compare treatments
```{r, message=FALSE, warning=FALSE}
spider_t2 <- which(M2_treatment$timepoint == "t2" & M2_treatment$Organism == "spider")
set.seed(222)
M2_anosim_spider_t2 <- anosim(M2_pufa_ma[spider_t2,], M2_treatment[spider_t2,]$treatment, distance = "euclidean")
summary(M2_anosim_spider_t2)

spider_t3 <- which(M2_treatment$timepoint == "t3" & M2_treatment$Organism == "spider")
set.seed(222)
M2_anosim_spider_t3 <- anosim(M2_pufa_ma[spider_t3,], M2_treatment[spider_t3,]$treatment, distance = "euclidean")
summary(M2_anosim_spider_t3)
```

### Spider within same treatment and start spiders: compare time points
```{r, message=FALSE, warning=FALSE}
M2_spider_leaves_time <- which(M2_treatment$Organism == "spider" & M2_treatment$treatment %in% c("leaves"))

set.seed(222)
M2_anosim_spider_leaves_time <- anosim(M2_pufa_ma[M2_spider_leaves_time,], M2_treatment[M2_spider_leaves_time,]$timepoint, distance = "euclidean")
summary(M2_anosim_spider_leaves_time)

M2_spider_oatmeal_time <- which(M2_treatment$Organism == "spider" & M2_treatment$treatment %in% c("oatmeal"))

set.seed(222)
M2_anosim_spider_oatmeal_time <- anosim(M2_pufa_ma[M2_spider_oatmeal_time,], M2_treatment[M2_spider_oatmeal_time,]$timepoint, distance = "euclidean")
summary(M2_anosim_spider_oatmeal_time)

M2_spider_fishfood_time <- which(M2_treatment$Organism == "spider" & M2_treatment$treatment %in% c("fishfood"))

set.seed(222)
M2_anosim_spider_fishfood_time <- anosim(M2_pufa_ma[M2_spider_fishfood_time,], M2_treatment[M2_spider_fishfood_time,]$timepoint, distance = "euclidean")
summary(M2_anosim_spider_fishfood_time)
```

### Chiro within same time point: compare treatments
```{r, message=FALSE, warning=FALSE}
chiro_t2 <- which(M2_treatment$timepoint == "t2" & M2_treatment$Organism == "chiro")
set.seed(222)
M2_anosim_chiro_t2 <- anosim(M2_pufa_ma[chiro_t2,], M2_treatment[chiro_t2,]$treatment, distance = "euclidean")
summary(M2_anosim_chiro_t2)

chiro_t3 <- which(M2_treatment$timepoint == "t3" & M2_treatment$Organism == "chiro")
set.seed(222)
M2_anosim_chiro_t3 <- anosim(M2_pufa_ma[chiro_t3,], M2_treatment[chiro_t3,]$treatment, distance = "euclidean")
summary(M2_anosim_chiro_t3)
```

### Food source within same time point: compare treatments
```{r, message=FALSE, warning=FALSE}
food_t2 <- which(M2_treatment$timepoint == "t2" & M2_treatment$Organism %in% c("leaves", "oatmeal", "fishfood"))

set.seed(222)
M2_anosim_food_t2 <- anosim(M2_pufa_ma[food_t2,], M2_treatment[food_t2,]$treatment, distance = "euclidean")
summary(M2_anosim_food_t2)

food_t3 <- which(M2_treatment$timepoint == "t3" & M2_treatment$Organism %in% c("leaves", "oatmeal", "fishfood"))

set.seed(222)
M2_anosim_food_t3 <- anosim(M2_pufa_ma[food_t3,], M2_treatment[food_t3,]$treatment, distance = "euclidean")
summary(M2_anosim_food_t3)
```

### Extract R and p-values
```{r, message=FALSE, warning=FALSE}
M2_anosim_name <- c("M2_anosim_spider_t2", "M2_anosim_spider_t3",
                    "M2_anosim_spider_leaves_time", "M2_anosim_spider_oatmeal_time",
                    "M2_anosim_spider_fishfood_time", "M2_anosim_chiro_t2",
                    "M2_anosim_chiro_t3", "M2_anosim_food_t2", "M2_anosim_food_t3")

M2_anosim_p <- rbind(M2_anosim_spider_t2$signif, M2_anosim_spider_t3$signif,
                 M2_anosim_spider_leaves_time$signif, M2_anosim_spider_oatmeal_time$signif,
                 M2_anosim_spider_fishfood_time$signif, M2_anosim_chiro_t2$signif,
                 M2_anosim_chiro_t3$signif, M2_anosim_food_t2$signif,
                 M2_anosim_food_t3$signif)

M2_anosim_R <- rbind(M2_anosim_spider_t2$statistic, M2_anosim_spider_t3$statistic,
                 M2_anosim_spider_leaves_time$statistic,
                 M2_anosim_spider_oatmeal_time$statistic,
                 M2_anosim_spider_fishfood_time$statistic, M2_anosim_chiro_t2$statistic,
                 M2_anosim_chiro_t3$statistic, M2_anosim_food_t2$statistic,
                 M2_anosim_food_t3$statistic)

M2_anosim <- as.data.frame(cbind(M2_anosim_name, M2_anosim_p, M2_anosim_R))

names(M2_anosim)[2] <- "p_value"
names(M2_anosim)[3] <- "R"
```

### Adjust p-values and save output
```{r, message=FALSE, warning=FALSE}
M2_anosim$p_value_adjusted <- p.adjust(M2_anosim$p_value, method = "hochberg")

#write.csv(M2_anosim, "M2_anosim.csv", row.names = FALSE)
```

# SIMPER

* SIMPER analysis, where treatment was significant in ANOSIM
* In first experiment treatment always significant
* In second experiment treatment significant för chiros and food sources, but not for spiders

## M1
### Spiders
```{r, message=FALSE, warning=FALSE}
M1_simper_spider_t4 <- simper(M1_pufa_ma[spider_t4,], 
                       group =  M1_treatment[spider_t4,]$treatment)

summary(M1_simper_spider_t4)


M1_simper_spider_t7 <- simper(M1_pufa_ma[spider_t7,], 
                       group =  M1_treatment[spider_t7,]$treatment)

summary(M1_simper_spider_t7)
```

### Chiro
```{r, message=FALSE, warning=FALSE}
M1_simper_chiro_t4 <- simper(M1_pufa_ma[chiro_t4,], 
                       group =  M1_treatment[chiro_t4,]$treatment)

summary(M1_simper_chiro_t4)

M1_simper_chiro_t7 <- simper(M1_pufa_ma[chiro_t7,], 
                       group =  M1_treatment[chiro_t7,]$treatment)

summary(M1_simper_chiro_t7)
```

### Food source
```{r, message=FALSE, warning=FALSE}
M1_simper_food_t4 <- simper(M1_pufa_ma[food_t4,], 
                       group =  M1_treatment[food_t4,]$treatment)

summary(M1_simper_food_t4)

M1_simper_food_t7 <- simper(M1_pufa_ma[food_t7,], 
                       group =  M1_treatment[food_t7,]$treatment)

summary(M1_simper_food_t7)
```

## M2
### Chiro
```{r, message=FALSE, warning=FALSE}
M2_simper_chiro_t2 <- simper(M2_pufa_ma[chiro_t2,], 
                       group =  M2_treatment[chiro_t2,]$treatment)

summary(M2_simper_chiro_t2)

M2_simper_chiro_t3 <- simper(M2_pufa_ma[chiro_t3,], 
                       group =  M2_treatment[chiro_t3,]$treatment)

summary(M2_simper_chiro_t3)
```

### Food source
```{r, message=FALSE, warning=FALSE}
M2_simper_food_t2 <- simper(M2_pufa_ma[food_t2,], 
                       group =  M2_treatment[food_t2,]$treatment)

summary(M2_simper_food_t2)

M2_simper_food_t3 <- simper(M2_pufa_ma[food_t3,], 
                       group =  M2_treatment[food_t3,]$treatment)

summary(M2_simper_food_t3)
```

## Save output
Get summary
```{r, message=FALSE, warning=FALSE}
M1_simper_spider_t4_summary <- summary(M1_simper_spider_t4)
M1_simper_spider_t7_summary <- summary(M1_simper_spider_t7)
M1_simper_chiro_t4_summary <- summary(M1_simper_chiro_t4)
M1_simper_chiro_t7_summary <- summary(M1_simper_chiro_t7)
M1_simper_food_t4_summary <- summary(M1_simper_food_t4)
M1_simper_food_t7_summary <- summary(M1_simper_food_t7)

M2_simper_chiro_t2_summary <- summary(M2_simper_chiro_t2)
M2_simper_chiro_t3_summary <- summary(M2_simper_chiro_t3)
M2_simper_food_t2_summary <- summary(M2_simper_food_t2)
M2_simper_food_t3_summary <- summary(M2_simper_food_t3)
```

Add time point to lists
```{r, message=FALSE, warning=FALSE}
  for (i in c(1:3))
{
M1_simper_spider_t4_summary[[i]]$Time_point <- "23 days"
  }

  for (i in c(1:3))
{
M1_simper_chiro_t4_summary[[i]]$Time_point <- "23 days"
  }

  for (i in c(1:3))
{
M1_simper_food_t4_summary[[i]]$Time_point <- "23 days"
  }

  for (i in c(1:3))
{
M1_simper_spider_t7_summary[[i]]$Time_point <- "44 days"
  }

  for (i in c(1:3))
{
M1_simper_chiro_t7_summary[[i]]$Time_point <- "44 days"
  }

  for (i in c(1:3))
{
M1_simper_food_t7_summary[[i]]$Time_point <- "44 days"
  }


  for (i in c(1:3))
{
M2_simper_chiro_t2_summary[[i]]$Time_point <- "14 days"
  }

  for (i in c(1:3))
{
M2_simper_food_t2_summary[[i]]$Time_point <- "14 days"
  }

  for (i in c(1:3))
{
M2_simper_chiro_t3_summary[[i]]$Time_point <- "21 days"
  }

  for (i in c(1:3))
{
M2_simper_food_t3_summary[[i]]$Time_point <- "21 days"
  }
```

Add trophic level to lists
```{r, message=FALSE, warning=FALSE}
  for (i in c(1:3))
{
M1_simper_spider_t4_summary[[i]]$Trophic_level <- "spiders"
  }

  for (i in c(1:3))
{
M1_simper_chiro_t4_summary[[i]]$Trophic_level <- "chironomids"
  }

  for (i in c(1:3))
{
M1_simper_food_t4_summary[[i]]$Trophic_level <- "basic food sources"
  }

  for (i in c(1:3))
{
M1_simper_spider_t7_summary[[i]]$Trophic_level <- "spiders"
  }

  for (i in c(1:3))
{
M1_simper_chiro_t7_summary[[i]]$Trophic_level <- "chironomids"
  }

  for (i in c(1:3))
{
M1_simper_food_t7_summary[[i]]$Trophic_level <- "basic food sources"
  }


  for (i in c(1:3))
{
M2_simper_chiro_t2_summary[[i]]$Trophic_level <- "chironomids"
  }

  for (i in c(1:3))
{
M2_simper_food_t2_summary[[i]]$Trophic_level <- "basic food sources"
  }

  for (i in c(1:3))
{
M2_simper_chiro_t3_summary[[i]]$Trophic_level <- "chironomids"
  }

  for (i in c(1:3))
{
M2_simper_food_t3_summary[[i]]$Trophic_level <- "basic food sources"
  }
```

Rownames as column
```{r, message=FALSE, warning=FALSE}
for (i in c(1:3))
{
M1_simper_spider_t4_summary[[i]]$PUFA <- rownames(M1_simper_spider_t4_summary[[i]]) 
  }

  for (i in c(1:3))
{
M1_simper_spider_t7_summary[[i]]$PUFA <- rownames(M1_simper_spider_t7_summary[[i]]) 
  }

  for (i in c(1:3))
{
M1_simper_chiro_t4_summary[[i]]$PUFA <- rownames(M1_simper_chiro_t4_summary[[i]]) 
  }

  for (i in c(1:3))
{
M1_simper_chiro_t7_summary[[i]]$PUFA <- rownames(M1_simper_chiro_t7_summary[[i]])
  }

  for (i in c(1:3))
{
M1_simper_food_t4_summary[[i]]$PUFA <- rownames(M1_simper_food_t4_summary[[i]]) 
  }

  for (i in c(1:3))
{
M1_simper_food_t7_summary[[i]]$PUFA <- rownames(M1_simper_food_t7_summary[[i]])
  }

  for (i in c(1:3))
{
M2_simper_chiro_t2_summary[[i]]$PUFA <- rownames(M2_simper_chiro_t2_summary[[i]]) 
  }

  for (i in c(1:3))
{
M2_simper_chiro_t3_summary[[i]]$PUFA <- rownames(M2_simper_chiro_t3_summary[[i]])
  }

  for (i in c(1:3))
{
M2_simper_food_t2_summary[[i]]$PUFA <- rownames(M2_simper_food_t2_summary[[i]]) 
  }

  for (i in c(1:3))
{
M2_simper_food_t3_summary[[i]]$PUFA <- rownames(M2_simper_food_t3_summary[[i]])
  }
```

Listnames as column
```{r, message=FALSE, warning=FALSE}
for (i in c(1:3))
{
M1_simper_spider_t4_summary[[i]]$Contrast <-names(M1_simper_spider_t4_summary[i]) 
  }

for (i in c(1:3))
{
M1_simper_spider_t7_summary[[i]]$Contrast <-names(M1_simper_spider_t7_summary[i]) 
}

  for (i in c(1:3))
{
M1_simper_chiro_t4_summary[[i]]$Contrast <- names(M1_simper_chiro_t4_summary[i]) 
  }

  for (i in c(1:3))
{
M1_simper_chiro_t7_summary[[i]]$Contrast <- names(M1_simper_chiro_t7_summary[i])
  }

  for (i in c(1:3))
{
M1_simper_food_t4_summary[[i]]$Contrast <- names(M1_simper_food_t4_summary[i]) 
  }

  for (i in c(1:3))
{
M1_simper_food_t7_summary[[i]]$Contrast <- names(M1_simper_food_t7_summary[i])
  }

  for (i in c(1:3))
{
M2_simper_chiro_t2_summary[[i]]$Contrast <- names(M2_simper_chiro_t2_summary[i]) 
  }

  for (i in c(1:3))
{
M2_simper_chiro_t3_summary[[i]]$Contrast <- names(M2_simper_chiro_t3_summary[i])
  }

  for (i in c(1:3))
{
M2_simper_food_t2_summary[[i]]$Contrast <- names(M2_simper_food_t2_summary[i]) 
  }

  for (i in c(1:3))
{
M2_simper_food_t3_summary[[i]]$Contrast <- names(M2_simper_food_t3_summary[i])
  }

```

Chose cumsum nearest to 70 % (0.7)
```{r, message=FALSE, warning=FALSE}
closest_greater <-  function(x, viable_numbers) {
  min(viable_numbers[viable_numbers >= x])
}

  for (i in c(1:3))
{
M1_simper_spider_t4_summary[[i]] <- filter(M1_simper_spider_t4_summary[[i]], 
                    cumsum <= closest_greater(0.7, M1_simper_spider_t4_summary[[i]]$cumsum))
  }

  for (i in c(1:3))
{
M1_simper_spider_t7_summary[[i]] <- filter(M1_simper_spider_t7_summary[[i]], 
                    cumsum <= closest_greater(0.7, M1_simper_spider_t7_summary[[i]]$cumsum))
  }

  for (i in c(1:3))
{
M1_simper_chiro_t4_summary[[i]] <- filter(M1_simper_chiro_t4_summary[[i]], 
                    cumsum <= closest_greater(0.7,M1_simper_chiro_t4_summary[[i]]$cumsum))
  }   
    
  for (i in c(1:3))
{
M1_simper_chiro_t7_summary[[i]] <- filter(M1_simper_chiro_t7_summary[[i]], 
                    cumsum <= closest_greater(0.7, M1_simper_chiro_t7_summary[[i]]$cumsum))
  }

  for (i in c(1:3))
{
M1_simper_food_t4_summary[[i]] <- filter(M1_simper_food_t4_summary[[i]], 
                    cumsum <= closest_greater(0.7, M1_simper_food_t4_summary[[i]]$cumsum))
  }

  for (i in c(1:3))
{
M1_simper_food_t7_summary[[i]] <- filter(M1_simper_food_t7_summary[[i]], 
                    cumsum <= closest_greater(0.7, M1_simper_food_t7_summary[[i]]$cumsum))
  }

for (i in c(1:3))
{
M2_simper_chiro_t2_summary[[i]] <- filter(M2_simper_chiro_t2_summary[[i]], 
                    cumsum <= closest_greater(0.7, M2_simper_chiro_t2_summary[[i]]$cumsum))
  }

  for (i in c(1:3))
{
M2_simper_chiro_t3_summary[[i]] <- filter(M2_simper_chiro_t3_summary[[i]], 
                    cumsum <= closest_greater(0.7, M2_simper_chiro_t3_summary[[i]]$cumsum))
  }

  for (i in c(1:3))
{
M2_simper_food_t2_summary[[i]] <- filter(M2_simper_food_t2_summary[[i]], 
                    cumsum <= closest_greater(0.7, M2_simper_food_t2_summary[[i]]$cumsum))
  }

  for (i in c(1:3))
{
M2_simper_food_t3_summary[[i]] <- filter(M2_simper_food_t3_summary[[i]], 
                    cumsum <= closest_greater(0.7, M2_simper_food_t3_summary[[i]]$cumsum))
  }
```

Lists to data frames
```{r, message=FALSE, warning=FALSE}
M1_simper_spider_t4_summary <- do.call(rbind, M1_simper_spider_t4_summary) 
M1_simper_spider_t7_summary <- do.call(rbind, M1_simper_spider_t7_summary) 
M1_simper_chiro_t4_summary <- do.call(rbind, M1_simper_chiro_t4_summary) 
M1_simper_chiro_t7_summary <- do.call(rbind, M1_simper_chiro_t7_summary)
M1_simper_food_t4_summary <- do.call(rbind, M1_simper_food_t4_summary) 
M1_simper_food_t7_summary <- do.call(rbind, M1_simper_food_t7_summary)

M2_simper_chiro_t2_summary <- do.call(rbind, M2_simper_chiro_t2_summary) 
M2_simper_chiro_t3_summary <- do.call(rbind, M2_simper_chiro_t3_summary)
M2_simper_food_t2_summary <- do.call(rbind, M2_simper_food_t2_summary) 
M2_simper_food_t3_summary <- do.call(rbind, M2_simper_food_t3_summary)
```

Bind summary data frames
```{r, message=FALSE, warning=FALSE}
simper_summary <- rbind(M1_simper_spider_t4_summary, M1_simper_spider_t7_summary,
                        M1_simper_chiro_t4_summary, M1_simper_chiro_t7_summary,
                        M1_simper_food_t4_summary, M1_simper_food_t7_summary,
                        M2_simper_chiro_t2_summary, M2_simper_chiro_t3_summary,
                        M2_simper_food_t2_summary, M2_simper_food_t3_summary)
```

Add column experiment
```{r, message=FALSE, warning=FALSE}
simper_summary$Experiment <- with(simper_summary, 
                                  ifelse(Time_point == "44 days", "1",
                                  ifelse(Time_point == "23 days", "1",
                                  2)))
```

Put column in other order
```{r, message=FALSE, warning=FALSE}
simper_summary <- select(simper_summary, 
                         c("Experiment", "Trophic_level", "Time_point",
                           "Contrast","PUFA", "average", "sd", 
                           "ratio", "ava", "avb", "cumsum" ))
```

Rename columns
```{r, message=FALSE, warning=FALSE}
colnames_simper_summary <- c("Experiment","Trophic level", "Time point", 
                             "Contrast", "PUFA", "Average", 
                           "Standard deviation", "Ratio", 
                           "Average a", "Average b", "Cumulative contribution")
names(simper_summary) <- colnames_simper_summary
```

Round 2 digits
```{r, message=FALSE, warning=FALSE}
simper_summary[,6:11] <- round(simper_summary[,6:11], digits = 2)
```

Rename PUFA
```{r, message=FALSE, warning=FALSE}

simper_summary$PUFA <- gsub("22:6n-3", "DHA", simper_summary$PUFA) # docosahexaenoic acid
  
simper_summary$PUFA <- gsub("20:5n-3", "EPA", simper_summary$PUFA) # eicosapentaenoic acid
simper_summary$PUFA <- gsub("20:4n-6", "ARA", simper_summary$PUFA) # arachidonic acid

simper_summary$PUFA <- gsub("18:3n-3", "ALA", simper_summary$PUFA) # alpha-linolenic acid
simper_summary$PUFA <- gsub("18:3n-6", "GLA", simper_summary$PUFA) # gamma-Linolenic acid

simper_summary$PUFA <- gsub("18:2n-6c", "LIN", simper_summary$PUFA) # Linoleic acid
simper_summary$PUFA <- gsub("18:2n-6t", "LLA", simper_summary$PUFA) # Linolelaidic acid
```

Save csv
```{r, message=FALSE, warning=FALSE}
#write.csv(simper_summary, file = "simper_summary.csv", row.names = FALSE)
```

Which PUFA contribute to differences
```{r, message=FALSE, warning=FALSE}
unique(simper_summary$PUFA)
```

Average a bigger average b
```{r, message=FALSE, warning=FALSE}
simper_summary$ava_bigger <- with(simper_summary, ifelse(`Average a` > `Average b`, "y", "n"))
```

Count which PUFA in contrast main contributer to difference
```{r, message=FALSE, warning=FALSE}
count_simper_summary <- count(simper_summary, c("Experiment", "Contrast",
                                                "ava_bigger","PUFA")) 

count_simper_summary_1 <- count(simper_summary, c("Experiment", "PUFA"))

View(filter(count_simper_summary, Experiment == 2))
```

# LM(M)
## Weekly endpoints

* fixed effects: fresh weight, growth and body condition
* random effect: spider_ID, because spiders are measured repeatedly


### M1
#### Body condition
```{r, message=FALSE, warning=FALSE}
M1_lmm_body_cond <-glmmTMB(body_cond ~ treatment * timepoint + (1 | spider_ID), 
                           family = "gaussian", data = M1_stat_weekly, na.action = na.omit)

summary(M1_lmm_body_cond)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_body <- fitted(M1_lmm_body_cond)
E_M1_body<- resid(M1_lmm_body_cond)

plot(x = F_M1_body, y = E_M1_body, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_body, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
M1_anova_body_cond <- Anova(M1_lmm_body_cond, type = c(2))

M1_anova_body_cond
```

Interaction significant so keep it

Plot interactions
```{r, message=FALSE, warning=FALSE}
plot(allEffects(M1_lmm_body_cond, x.var = "timepoint"))
```

#### Fresh weight
```{r, message=FALSE, warning=FALSE}
M1_lmm_fresh_weight <-glmmTMB(weight_spider_g ~ treatment * timepoint + (1 | spider_ID), 
                        family = "gaussian", data = M1_stat_weekly, na.action = na.omit,
                        control = glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

summary(M1_lmm_fresh_weight)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_fresh_weight <- fitted(M1_lmm_fresh_weight)
E_M1_fresh_weight<- resid(M1_lmm_fresh_weight)

plot(x = F_M1_fresh_weight, y = E_M1_fresh_weight, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_fresh_weight, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
M1_anova_fresh_weight <- Anova(M1_lmm_fresh_weight, type = c(2))

M1_anova_fresh_weight
```

Interaction significant so keep it

Plot interactions
```{r, message=FALSE, warning=FALSE}
plot(allEffects(M1_lmm_fresh_weight, x.var = "timepoint"))

plot(allEffects(M1_lmm_fresh_weight, x.var = "treatment"))
```

#### Growth rate 
```{r, message=FALSE, warning=FALSE}
M1_lmm_growth_rate <-glmmTMB(growth_rate ~ treatment * timepoint 
                               + (1 | spider_ID), 
                        family = "gaussian", 
                        data = M1_stat_weekly[M1_stat_weekly$timepoint != "t0",], 
                        na.action = na.omit,
                        control = 
                          glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

summary(M1_lmm_growth_rate)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_growth_rate <- fitted(M1_lmm_growth_rate)
E_M1_growth_rate<- resid(M1_lmm_growth_rate)

plot(x = F_M1_growth_rate, y = E_M1_growth_rate, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_growth_rate, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
M1_anova_growth_rate <- Anova(M1_lmm_growth_rate, type = c(2))

M1_anova_growth_rate
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M1_lmm_growth_rate_1 <-glmmTMB(growth_rate ~ treatment + timepoint 
                               + (1 | spider_ID), 
                        family = "gaussian", 
                        data = M1_stat_weekly[M1_stat_weekly$timepoint != "t0",], 
                        na.action = na.omit,
                        control = 
                          glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

summary(M1_lmm_growth_rate_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_growth_rate <- fitted(M1_lmm_growth_rate_1)
E_M1_growth_rate<- resid(M1_lmm_growth_rate_1)

plot(x = F_M1_growth_rate, y = E_M1_growth_rate, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_growth_rate, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
M1_anova_growth_rate <- Anova(M1_lmm_growth_rate_1, type = c(2))

M1_anova_growth_rate
```

### M2 
#### Body condition
```{r, message=FALSE, warning=FALSE}
M2_lmm_body_cond <-glmmTMB(body_cond ~ treatment * timepoint + (1 | spider_ID), 
                           family = "gaussian", data = M2_stat_weekly, na.action = na.omit)

summary(M2_lmm_body_cond)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_body <- fitted(M2_lmm_body_cond)
E_M2_body<- resid(M2_lmm_body_cond)

plot(x = F_M2_body, y = E_M2_body, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_body, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
Anova(M2_lmm_body_cond, type = c(2))
```

Interaction not significant so model without interaction
```{r, message=FALSE, warning=FALSE}
M2_lmm_body_cond_1 <-glmmTMB(body_cond ~ treatment + timepoint + (1 | spider_ID), 
                           family = "gaussian", data = M2_stat_weekly, na.action = na.omit)

summary(M2_lmm_body_cond_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_body <- fitted(M2_lmm_body_cond_1)
E_M2_body<- resid(M2_lmm_body_cond_1)

plot(x = F_M2_body, y = E_M2_body, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_body, nrep = 8) # normality okay
```

Test effects of explanatory variables with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
M2_anova_body_cond <- Anova(M2_lmm_body_cond_1, type = c(2))
M2_anova_body_cond
```

#### Fresh weight
```{r, message=FALSE, warning=FALSE}
M2_lmm_fresh_weight <-glmmTMB(weight_spider_g ~ treatment * timepoint + (1 | spider_ID), 
                        family = "gaussian", data = M2_stat_weekly, na.action = na.omit)

summary(M2_lmm_fresh_weight)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_fresh_weight <- fitted(M2_lmm_fresh_weight)
E_M2_fresh_weight<- resid(M2_lmm_fresh_weight)

plot(x = F_M2_fresh_weight, y = E_M2_fresh_weight, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_fresh_weight, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
Anova(M2_lmm_fresh_weight, type = c(2))
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M2_lmm_fresh_weight_1 <-glmmTMB(weight_spider_g ~ treatment + timepoint + (1 | spider_ID), 
                        family = "gaussian", data = M2_stat_weekly, na.action = na.omit)

summary(M2_lmm_fresh_weight_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_fresh_weight <- fitted(M2_lmm_fresh_weight_1)
E_M2_growth<- resid(M2_lmm_fresh_weight_1)

plot(x = F_M2_fresh_weight, y = E_M2_growth, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_growth, nrep = 8) # normality okay
```

Test effects of explanatory variables with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
M2_anova_fresh_weight <- Anova(M2_lmm_fresh_weight_1, type = c(2))
M2_anova_fresh_weight
```

#### Growth rate
```{r, message=FALSE, warning=FALSE}
M2_lmm_growth_rate <-glmmTMB(growth_rate ~ treatment * timepoint 
                               + (1 | spider_ID), 
                        family = "gaussian", 
                        data = M2_stat_weekly[M2_stat_weekly$timepoint != "t0",], 
                        na.action = na.omit,
                        control = 
                          glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

summary(M2_lmm_growth_rate)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_growth_rate <- fitted(M2_lmm_growth_rate)
E_M2_growth_rate<- resid(M2_lmm_growth_rate)

plot(x = F_M2_growth_rate, y = E_M2_growth_rate, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_growth_rate, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
M2_anova_growth_rate <- Anova(M2_lmm_growth_rate, type = c(2))

M2_anova_growth_rate
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M2_lmm_growth_rate_1 <-glmmTMB(growth_rate ~ treatment + timepoint 
                               + (1 | spider_ID), 
                        family = "gaussian", 
                        data = M2_stat_weekly[M2_stat_weekly$timepoint != "t0",], 
                        na.action = na.omit,
                        control = 
                          glmmTMBControl(optCtrl=list(iter.max=1e3,eval.max=1e3)))

summary(M2_lmm_growth_rate_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_growth_rate <- fitted(M2_lmm_growth_rate_1)
E_M2_growth_rate<- resid(M2_lmm_growth_rate_1)

plot(x = F_M2_growth_rate, y = E_M2_growth_rate, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_growth_rate, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using Wald chisquare tests.
```{r, message=FALSE, warning=FALSE}
M2_anova_growth_rate <- Anova(M2_lmm_growth_rate_1, type = c(2))

M2_anova_growth_rate
```

## Deadly endpoints

* spider_ID not used as random effect, because every spider measured only once

### M1
#### Dry weight
```{r, message=FALSE, warning=FALSE}
M1_lm_dry_weight <- lm(dry_weight_mg ~ treatment * timepoint, data = M1_stat_deadly,
                       na.action = na.omit)

summary(M1_lm_dry_weight)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_dry_weight <- fitted(M1_lm_dry_weight)
E_M1_dry_weight <- resid(M1_lm_dry_weight)

plot(x = F_M1_dry_weight, y = E_M1_dry_weight, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_dry_weight, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
Anova(M1_lm_dry_weight, type = c(2))
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M1_lm_dry_weight_1 <- lm(dry_weight_mg ~ treatment + timepoint, data = M1_stat_deadly,
                       na.action = na.omit)

summary(M1_lm_dry_weight_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_dry_weight <- fitted(M1_lm_dry_weight_1)
E_M1_dry_weight <- resid(M1_lm_dry_weight_1)

plot(x = F_M1_dry_weight, y = E_M1_dry_weight, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_dry_weight, nrep = 8) # normality okay
```

Test effects of explanatory variables with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
M1_anova_dry_weight <- Anova(M1_lm_dry_weight_1, type = c(2))

M1_anova_dry_weight
```

#### Ratio encapsulated, uncapsulated area
```{r, message=FALSE, warning=FALSE}
M1_lm_ratio_encapsulated <- lm(mean_R_e_ne ~ treatment * timepoint, data = M1_stat_deadly,
                                na.action = na.omit)

summary(M1_lm_ratio_encapsulated)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_ratio_encapsulated <- fitted(M1_lm_ratio_encapsulated)
E_M1_ratio_encapsulated <- resid(M1_lm_ratio_encapsulated)

plot(x = F_M1_ratio_encapsulated, y = E_M1_ratio_encapsulated, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_ratio_encapsulated, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
Anova(M1_lm_ratio_encapsulated, type = c(2))
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M1_lm_ratio_encapsulated_1 <- lm(mean_R_e_ne ~ treatment + timepoint, 
                                 data = M1_stat_deadly,
                                na.action = na.omit)

summary(M1_lm_ratio_encapsulated_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_ratio_encapsulated <- fitted(M1_lm_ratio_encapsulated_1)
E_M1_ratio_encapsulated <- resid(M1_lm_ratio_encapsulated_1)

plot(x = F_M1_ratio_encapsulated, y = E_M1_ratio_encapsulated, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_ratio_encapsulated, nrep = 8) # normality okay
```

Test effects of explanatory variables with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
M1_anova_ratio_encapsulated <- Anova(M1_lm_ratio_encapsulated_1, type = c(2))
M1_anova_ratio_encapsulated
```

#### Area encapsulated
```{r, message=FALSE, warning=FALSE}
M1_lm_area_encapsulated <- lm(mean_area_encapsulation ~ treatment * timepoint, 
                              data = M1_stat_deadly, na.action = na.omit)

summary(M1_lm_area_encapsulated)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_area_encapsulated <- fitted(M1_lm_area_encapsulated)
E_M1_area_encapsulated <- resid(M1_lm_area_encapsulated)

plot(x = F_M1_area_encapsulated, y = E_M1_area_encapsulated, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_area_encapsulated, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
Anova(M1_lm_area_encapsulated, type = c(2))
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M1_lm_area_encapsulated_1 <- lm(mean_area_encapsulation ~ treatment + timepoint, 
                              data = M1_stat_deadly, na.action = na.omit)

summary(M1_lm_area_encapsulated_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M1_area_encapsulated <- fitted(M1_lm_area_encapsulated_1)
E_M1_area_encapsulated <- resid(M1_lm_area_encapsulated_1)

plot(x = F_M1_area_encapsulated, y = E_M1_area_encapsulated, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M1_area_encapsulated, nrep = 8) # normality okay
```

Test effects of explanatory variables with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
M1_anova_area_encapsulated <- Anova(M1_lm_area_encapsulated_1, type = c(2))
M1_anova_area_encapsulated
```

### M2
#### Dry weight
```{r, message=FALSE, warning=FALSE}
M2_lm_dry_weight <- lm(dry_weight_mg ~ treatment * timepoint, data = M2_stat_deadly,
                       na.action = na.omit)

summary(M2_lm_dry_weight)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_dry_weight <- fitted(M2_lm_dry_weight)
E_M2_dry_weight <- resid(M2_lm_dry_weight)

plot(x = F_M2_dry_weight, y = E_M2_dry_weight, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_dry_weight, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
Anova(M2_lm_dry_weight, type = c(2))
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M2_lm_dry_weight_1 <- lm(dry_weight_mg ~ treatment + timepoint, data = M2_stat_deadly,
                       na.action = na.omit)

summary(M2_lm_dry_weight_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_dry_weight <- fitted(M2_lm_dry_weight_1)
E_M2_dry_weight <- resid(M2_lm_dry_weight_1)

plot(x = F_M2_dry_weight, y = E_M2_dry_weight, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_dry_weight, nrep = 8) # normality okay
```

Test effects of explanatory variables with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
M2_anova_dry_weight <- Anova(M2_lm_dry_weight_1, type = c(2))
M2_anova_dry_weight
```

#### Ratio encapsulated, uncapsulated area
```{r, message=FALSE, warning=FALSE}
M2_lm_ratio_encapsulated <- lm(mean_R_e_ne ~ treatment * timepoint, data = M2_stat_deadly,
                               na.action = na.omit)

summary(M2_lm_ratio_encapsulated)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_ratio_encapsulated <- fitted(M2_lm_ratio_encapsulated)
E_M2_ratio_encapsulated <- resid(M2_lm_ratio_encapsulated)

plot(x = F_M2_ratio_encapsulated, y = E_M2_ratio_encapsulated, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance ok

qreference(E_M2_ratio_encapsulated, nrep = 8) # normality ok
```

Test effects of explanatory variables and their interactions with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
Anova(M2_lm_ratio_encapsulated, type = c(2))
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M2_lm_ratio_encapsulated_1 <- lm(mean_R_e_ne ~ treatment + timepoint, data = M2_stat_deadly,
                               na.action = na.omit)

summary(M2_lm_ratio_encapsulated_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_ratio_encapsulated <- fitted(M2_lm_ratio_encapsulated_1)
E_M2_ratio_encapsulated <- resid(M2_lm_ratio_encapsulated_1)

plot(x = F_M2_ratio_encapsulated, y = E_M2_ratio_encapsulated, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance ok

qreference(E_M2_ratio_encapsulated, nrep = 8) # normality ok
```

Test effects of explanatory variables with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
M2_anova_ratio_encapsulated <- Anova(M2_lm_ratio_encapsulated_1, type = c(2))
M2_anova_ratio_encapsulated
```

#### Area encapsulated
```{r, message=FALSE, warning=FALSE}
M2_lm_area_encapsulated <- lm(mean_area_encapsulation ~ treatment * timepoint,
                              data = M2_stat_deadly, na.action = na.omit)

summary(M2_lm_area_encapsulated)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_area_encapsulated <- fitted(M2_lm_area_encapsulated)
E_M2_area_encapsulated <- resid(M2_lm_area_encapsulated)

plot(x = F_M2_area_encapsulated, y = E_M2_area_encapsulated, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_area_encapsulated, nrep = 8) # normality okay
```

Test effects of explanatory variables and their interactions with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
Anova(M2_lm_area_encapsulated, type = c(2))
```

Interaction not significant, so model without interaction
```{r, message=FALSE, warning=FALSE}
M2_lm_area_encapsulated_1 <- lm(mean_area_encapsulation ~ treatment + timepoint,
                              data = M2_stat_deadly, na.action = na.omit)

summary(M2_lm_area_encapsulated_1)
```

Check normality and variance
```{r, message=FALSE, warning=FALSE}
F_M2_area_encapsulated <- fitted(M2_lm_area_encapsulated_1)
E_M2_area_encapsulated <- resid(M2_lm_area_encapsulated_1)

plot(x = F_M2_area_encapsulated, y = E_M2_area_encapsulated, xlab = "Fitted values", ylab = "Pearson residuals")
abline(h = 0, lty = 2) # variance okay

qreference(E_M2_area_encapsulated, nrep = 8) # normality okay
```

Test effects of explanatory variables with type II ANOVA using F-tests.
```{r, message=FALSE, warning=FALSE}
M2_anova_area_encapsulated <- Anova(M2_lm_area_encapsulated_1, type = c(2))
M2_anova_area_encapsulated
```

## Export results

Variable names
```{r, message=FALSE, warning=FALSE}
lmm_variable <- c("treatment", "timepoint", "treatment:timepoint")
lmm_variable_1 <- c("treatment", "timepoint")

lm_variable <- c("treatment", "timepoint", "residuals")
```

Test statistics for LMM
```{r, message=FALSE, warning=FALSE}
lmm_anova_test_stat <- as.data.frame(cbind(M1_anova_fresh_weight$Chisq,
                                           M1_anova_body_cond$Chisq))
lmm_anova_test_stat$expl_variable <- lmm_variable
lmm_anova_test_stat$variable <- "Chisq"

lmm_anova_test_stat_1 <- as.data.frame(cbind(M1_anova_growth_rate$Chisq,
                                             M2_anova_fresh_weight$Chisq,
                                             M2_anova_body_cond$Chisq,
                                             M2_anova_growth_rate$Chisq))
lmm_anova_test_stat_1$expl_variable <- lmm_variable_1
lmm_anova_test_stat_1$variable <- "Chisq"
```

Test statistic for LM
```{r, message=FALSE, warning=FALSE}
lm_anova_test_stat <- as.data.frame(cbind(M1_anova_dry_weight$'F value', 
                                          M1_anova_ratio_encapsulated$'F value', 
                                          M1_anova_area_encapsulated$'F value',
                                          M2_anova_dry_weight$'F value',
                                          M2_anova_ratio_encapsulated$'F value', 
                                          M2_anova_area_encapsulated$'F value'))
lm_anova_test_stat$expl_variable <- lm_variable
lm_anova_test_stat$variable <- "F value"
```

p-values
```{r, message=FALSE, warning=FALSE}
lmm_anova_p_value <- as.data.frame(cbind(M1_anova_fresh_weight$'Pr(>Chisq)',
                                         M1_anova_body_cond$'Pr(>Chisq)'))
lmm_anova_p_value$expl_variable <- lmm_variable
lmm_anova_p_value$variable <- "Pr(>Chisq)"

lmm_anova_p_value_1 <- as.data.frame(cbind(M1_anova_growth_rate$'Pr(>Chisq)',
                                           M2_anova_fresh_weight$'Pr(>Chisq)',
                                           M2_anova_body_cond$'Pr(>Chisq)',
                                           M2_anova_growth_rate$'Pr(>Chisq)'))
lmm_anova_p_value_1$expl_variable <- lmm_variable_1
lmm_anova_p_value_1$variable <- "Pr(>Chisq)"

lm_anova_p_value <- as.data.frame(cbind(M1_anova_dry_weight$'Pr(>F)',
                          M1_anova_ratio_encapsulated$'Pr(>F)',
                          M1_anova_area_encapsulated$'Pr(>F)', 
                          M2_anova_dry_weight$'Pr(>F)',
                          M2_anova_ratio_encapsulated$'Pr(>F)',
                          M2_anova_area_encapsulated$'Pr(>F)'))
lm_anova_p_value$expl_variable <- lm_variable
lm_anova_p_value$variable <- "Pr(>F)"
```

Degrees of freedom
```{r, message=FALSE, warning=FALSE}
lmm_anova_df <- as.data.frame(cbind(M1_anova_fresh_weight$Df, 
                                    M1_anova_body_cond$Df))
lmm_anova_df$expl_variable <- lmm_variable
lmm_anova_df$variable <- "Df"

lmm_anova_df_1 <- as.data.frame(cbind(M1_anova_growth_rate$Df,
                                      M2_anova_fresh_weight$Df,
                                      M2_anova_body_cond$Df,
                                      M2_anova_growth_rate$Df))
lmm_anova_df_1$expl_variable <- lmm_variable_1
lmm_anova_df_1$variable <- "Df"

lm_anova_df <- as.data.frame(cbind(M1_anova_dry_weight$Df,
                                   M1_anova_ratio_encapsulated$Df,
                                   M1_anova_area_encapsulated$Df,
                                   M2_anova_dry_weight$Df,
                                   M2_anova_ratio_encapsulated$Df,
                                   M2_anova_area_encapsulated$Df))
lm_anova_df$expl_variable <- lm_variable
lm_anova_df$variable <- "Df"
```

Save LMM 
```{r, message=FALSE, warning=FALSE}
lmm_anova <- rbind(lmm_anova_test_stat, lmm_anova_p_value, lmm_anova_df)
names(lmm_anova)[1:2] <- c("M1_anova_fresh_weight", "M1_anova_body_cond")

lmm_anova <- reshape2::melt(lmm_anova, id.vars = c("expl_variable", "variable"),
                            variable.name = "model")
lmm_anova <- reshape2::dcast(lmm_anova,  expl_variable ~ model + variable,
                             value.var = "value")

lmm_anova_1 <- rbind(lmm_anova_test_stat_1, lmm_anova_p_value_1, lmm_anova_df_1)
names(lmm_anova_1)[1:4] <- c("M1_anova_growth_rate", 
                             "M2_anova_fresh_weight",
                             "M2_anova_body_cond",
                             "M2_anova_growth_rate")
lmm_anova_1 <- reshape2::melt(lmm_anova_1, id.vars = c("expl_variable", "variable"), variable.name = "model")
lmm_anova_1 <- reshape2::dcast(lmm_anova_1,  expl_variable ~ model + variable, value.var = "value")

#write.csv(lmm_anova, file = "lmm_anova.csv", row.names = FALSE)
#write.csv(lmm_anova_1, file = "lmm_anova_1.csv", row.names = FALSE)
```

Save LM
```{r, message=FALSE, warning=FALSE}
lm_anova <- rbind(lm_anova_test_stat, lm_anova_p_value, lm_anova_df)
names(lm_anova)[1:6] <- c("M1_anova_dry_weight", 
                          "M1_anova_ratio_encapsulated",
                          "M1_anova_area_encapsulated", 
                          "M2_anova_dry_weight",
                          "M2_anova_ratio_encapsulated", 
                          "M2_anova_area_encapsulated")
lm_anova <- reshape2::melt(lm_anova, id.vars = c("expl_variable", "variable"), variable.name = "model")
lm_anova <- reshape2::dcast(lm_anova,  expl_variable ~ model + variable, value.var = "value")
#write.csv(lm_anova, file = "lm_anova.csv", row.names = FALSE)
```

# Plot weekly and deadly endpoints

## Preparation

Get mean and sd weekly endpoints
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly_mean <- select(M1_stat_weekly, 
                              -c("Organism", "experiment", "experiment_ID", "spider_ID", 
                                 "date_endpoint"))

M1_stat_weekly_mean <- reshape2::melt(M1_stat_weekly_mean, id.vars = c("treatment", "timepoint"))

M1_stat_weekly_mean <-  ddply(M1_stat_weekly_mean, c("treatment", "timepoint",
                                                     "variable"), summarise,
                              mean = mean(value, na.rm = TRUE),
                              sd = sd(value, na.rm = TRUE))

M2_stat_weekly_mean <- select(M2_stat_weekly, 
                              -c("Organism", "experiment", "experiment_ID", "spider_ID", 
                                 "date_endpoint"))

M2_stat_weekly_mean <- reshape2::melt(M2_stat_weekly_mean, id.vars = c("treatment", "timepoint"))

M2_stat_weekly_mean <-  ddply(M2_stat_weekly_mean, c("treatment", "timepoint",
                                                     "variable"), summarise,
                              mean = mean(value, na.rm = TRUE),
                              sd = sd(value, na.rm = TRUE))
```

Get mean and sd deadly endpoints
```{r, message=FALSE, warning=FALSE}
M1_stat_deadly_mean <- select(M1_stat_deadly, 
                              c("treatment", "timepoint", "mean_R_e_ne",
                                "mean_area_encapsulation", "dry_weight_mg"))

M1_stat_deadly_mean <- reshape2::melt(M1_stat_deadly_mean, id.vars = c("treatment", "timepoint"))

M1_stat_deadly_mean <-  ddply(M1_stat_deadly_mean, c("treatment", "timepoint",
                                                     "variable"), summarise,
                              mean = mean(value, na.rm = TRUE),
                              sd = sd(value, na.rm = TRUE))

M2_stat_deadly_mean <- select(M2_stat_deadly, 
                              c("treatment", "timepoint", "mean_R_e_ne",
                                "mean_area_encapsulation", "dry_weight_mg"))

M2_stat_deadly_mean <- reshape2::melt(M2_stat_deadly_mean, id.vars = c("treatment", "timepoint"))

M2_stat_deadly_mean <-  ddply(M2_stat_deadly_mean, c("treatment", "timepoint",
                                                     "variable"), summarise,
                              mean = mean(value, na.rm = TRUE),
                              sd = sd(value, na.rm = TRUE))
```

Rename fish food
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly_mean$treatment <- gsub("fishfood", "fish food",
                                      M1_stat_weekly_mean$treatment)
M1_stat_deadly_mean$treatment <- gsub("fishfood", "fish food",
                                      M1_stat_deadly_mean$treatment)

M2_stat_weekly_mean$treatment <- gsub("fishfood", "fish food",
                                      M2_stat_weekly_mean$treatment)
M2_stat_deadly_mean$treatment <- gsub("fishfood", "fish food",
                                      M2_stat_deadly_mean$treatment)
```

Rename time points
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly_mean$timepoint <- gsub("t0", "0 days",
                                      M1_stat_weekly_mean$timepoint)
M1_stat_weekly_mean$timepoint <- gsub("t2", "9 days",
                                      M1_stat_weekly_mean$timepoint)
M1_stat_weekly_mean$timepoint <- gsub("t3", "16 days",
                                      M1_stat_weekly_mean$timepoint)
M1_stat_weekly_mean$timepoint <- gsub("t4", "23 days",
                                      M1_stat_weekly_mean$timepoint)
M1_stat_weekly_mean$timepoint <- gsub("t5", "30 days",
                                      M1_stat_weekly_mean$timepoint)
M1_stat_weekly_mean$timepoint <- gsub("t6", "37 days",
                                      M1_stat_weekly_mean$timepoint)
M1_stat_weekly_mean$timepoint <- gsub("t7", "44 days",
                                      M1_stat_weekly_mean$timepoint)

M1_stat_deadly_mean$timepoint <- gsub("t0", "0 days",
                                      M1_stat_deadly_mean$timepoint)
M1_stat_deadly_mean$timepoint <- gsub("t4", "23 days",
                                      M1_stat_deadly_mean$timepoint)
M1_stat_deadly_mean$timepoint <- gsub("t7", "44 days",
                                      M1_stat_deadly_mean$timepoint)

M2_stat_weekly_mean$timepoint <- gsub("t0", "0 days", M2_stat_weekly_mean$timepoint)
M2_stat_weekly_mean$timepoint <- gsub("t1", "7 days", M2_stat_weekly_mean$timepoint)
M2_stat_weekly_mean$timepoint <- gsub("t2", "14 days", M2_stat_weekly_mean$timepoint)
M2_stat_weekly_mean$timepoint <- gsub("t3", "21 days", M2_stat_weekly_mean$timepoint)

M2_stat_deadly_mean$timepoint <- gsub("t0", "0 days", M2_stat_deadly_mean$timepoint)
M2_stat_deadly_mean$timepoint <- gsub("t2", "14 days", M2_stat_deadly_mean$timepoint)
M2_stat_deadly_mean$timepoint <- gsub("t3", "21 days", M2_stat_deadly_mean$timepoint)
```

Reorder factors
```{r, message=FALSE, warning=FALSE}
M1_stat_weekly_mean$treatment <- factor(M1_stat_weekly_mean$treatment, 
                                        levels = c("algae", "oatmeal", "fish food"))
M1_stat_deadly_mean$treatment <- factor(M1_stat_deadly_mean$treatment, 
                                        levels = c("algae", "oatmeal", "fish food"))

M1_stat_weekly_mean$timepoint <- factor(M1_stat_weekly_mean$timepoint, 
                                        levels = c("0 days", "9 days", "16 days", 
                                                   "23 days", "30 days", "37 days",
                                                   "44 days"))
M1_stat_deadly_mean$timepoint <- factor(M1_stat_deadly_mean$timepoint, 
                                        levels = c("0 days", "23 days", "44 days"))


M2_stat_weekly_mean$treatment <- factor(M2_stat_weekly_mean$treatment, 
                                        levels = c("leaves", "oatmeal", "fish food"))
M2_stat_deadly_mean$treatment <- factor(M2_stat_deadly_mean$treatment, 
                                        levels = c("leaves", "oatmeal", "fish food"))

M2_stat_weekly_mean$timepoint <- factor(M2_stat_weekly_mean$timepoint, 
                                        levels = c("0 days", "7 days", "14 days", 
                                                   "21 days"))
M2_stat_deadly_mean$test <- factor(M2_stat_deadly_mean$timepoint, 
                                        levels = c("0 days", "14 days", "21 days"))
```

## Plot
### Weekly endpoints
```{r, message=FALSE, warning=FALSE}
M1_fresh_weight_plot <- ggplot() +
  theme_classic() + 
  theme(legend.position="none",
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "Experiment 1", 
       x=expression(),
       y=expression(~ Fresh  ~ weight ~ "(" ~ g ~ ")")) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3")) +
  facet_grid(~ timepoint) +
  geom_point(data = M1_stat_weekly_mean[M1_stat_weekly_mean$variable == "weight_spider_g",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M1_stat_weekly_mean[M1_stat_weekly_mean$variable == "weight_spider_g",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
  ylim(0, 0.05)

M1_fresh_weight_plot

M1_body_plot <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="bottom", 
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "", 
       x=expression(),
       y=expression(~ body ~ condition)) +
  guides(color = guide_legend(override.aes = list(size = 5, shape = c(21,21,NA)))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3"),
                     labels = c("algae", "oatmeal", "")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3"),
                    labels = c("algae", "oatmeal", "")) +
  facet_grid(~ timepoint) +
  geom_point(data = M1_stat_weekly_mean[M1_stat_weekly_mean$variable == "body_cond",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M1_stat_weekly_mean[M1_stat_weekly_mean$variable == "body_cond",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
    ylim(0, 1.8)

M1_body_plot

M1_growth_rate_plot <- ggplot() +
  theme_classic() + 
  theme(legend.position="none",
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "", 
       x=expression(),
       y=expression(~ Growth  ~ rate ~ "("~g~ ~week^-1~")")) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3")) +
  facet_grid(~ timepoint) +
  geom_point(data = M1_stat_weekly_mean[M1_stat_weekly_mean$variable == "growth_rate" & M1_stat_weekly_mean$timepoint != "0 days",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M1_stat_weekly_mean[M1_stat_weekly_mean$variable == "growth_rate" & M1_stat_weekly_mean$timepoint != "0 days",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
  ylim(-0.3, 0.3)

M1_growth_rate_plot

M2_fresh_weight_plot <- ggplot() +
  theme_classic() + 
  theme(legend.position="none",
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "Experiment 2", 
       x=expression(),
       y=expression()) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3")) +
  facet_grid(~ timepoint) +
  geom_point(data = M2_stat_weekly_mean[M2_stat_weekly_mean$variable == "weight_spider_g",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M2_stat_weekly_mean[M2_stat_weekly_mean$variable == "weight_spider_g",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
  ylim(0, 0.05)

M2_fresh_weight_plot

M2_body_plot <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_blank(), legend.position="bottom", 
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "", 
       x=expression(),
       y=expression()) +
  guides(color = guide_legend(override.aes = list(size = 5, 
                                                  shape = c(21,NA,21)))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3"),
                     labels = c("leaves", "", "fishfood")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3"),
                    labels = c("leaves", "", "fishfood")) +
  facet_grid(~ timepoint) +
  geom_point(data = M2_stat_weekly_mean[M2_stat_weekly_mean$variable == "body_cond",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M2_stat_weekly_mean[M2_stat_weekly_mean$variable == "body_cond",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
    ylim(0, 1.8)

M2_body_plot

M2_growth_rate_plot <- ggplot() +
  theme_classic() + 
  theme(legend.position="none",
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "", 
       x=expression(),
       y=expression()) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3")) +
  facet_grid(~ timepoint) +
  geom_point(data = M2_stat_weekly_mean[M2_stat_weekly_mean$variable == "growth_rate" & M2_stat_weekly_mean$timepoint != "0 days",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M2_stat_weekly_mean[M2_stat_weekly_mean$variable == "growth_rate" & M2_stat_weekly_mean$timepoint != "0 days",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
  ylim(-0.3, 0.3)

M2_growth_rate_plot


weekly_endpoints_plot <- ggarrange(M1_fresh_weight_plot, M2_fresh_weight_plot,
                                   M1_growth_rate_plot, M2_growth_rate_plot,
                                   M1_body_plot, M2_body_plot,
                  labels = c("A", "B", "C", "D", "E", "F"),
                  ncol = 2, nrow = 3)

#png(filename="weekly_endpoints_plot.png", width=10, height=15, units="in", res=300)
weekly_endpoints_plot
#dev.off()
```

### Deadly endpoints
```{r, message=FALSE, warning=FALSE}
M1_weight_plot <- ggplot() +
  theme_classic() + 
  theme(legend.position="none",
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "Experiment 1", 
       x=expression(),
       y=expression(~ dry ~ weight ~ "(" ~ mg ~ ")")) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3")) +
  facet_grid(~ timepoint) +
  geom_point(data = M1_stat_deadly_mean[M1_stat_deadly_mean$variable == "dry_weight_mg",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M1_stat_deadly_mean[M1_stat_deadly_mean$variable == "dry_weight_mg",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
 ylim(0, 10)

M1_weight_plot

M1_R_e_ne_plot <- ggplot() +
  theme_classic() + 
  theme(legend.position="none",
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "", 
       x=expression(),
       y=expression(~ proportional ~ encapsulation  ~ area)) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3")) +
  facet_grid(~ timepoint) +
  geom_point(data = M1_stat_deadly_mean[M1_stat_deadly_mean$variable == "mean_R_e_ne",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M1_stat_deadly_mean[M1_stat_deadly_mean$variable == "mean_R_e_ne",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
    ylim(0, 2.5)

M1_R_e_ne_plot

M1_area_encapsulation_plot <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="bottom", 
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "", 
       x=expression(),
       y=expression(~ area ~ of ~ encapsulation ~ "(" ~ mm^2 ~ ")")) +
  guides(color = guide_legend(override.aes = list(size = 5, shape = c(21,21,NA)))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3"),
                     labels = c("algae", "oatmeal", "")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3"),
                    labels = c("algae", "oatmeal", "")) +
  facet_grid(~ timepoint) +
  geom_point(data = M1_stat_deadly_mean[M1_stat_deadly_mean$variable == "mean_area_encapsulation",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M1_stat_deadly_mean[M1_stat_deadly_mean$variable == "mean_area_encapsulation",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
   ylim(-0.03, 0.09)
  

M1_area_encapsulation_plot

M2_weight_plot <- ggplot() +
  theme_classic() + 
  theme(legend.position="none",
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "Experiment 2", 
       x=expression(),
       y=expression()) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3")) +
  facet_grid(~ timepoint) +
  geom_point(data = M2_stat_deadly_mean[M2_stat_deadly_mean$variable == "dry_weight_mg",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M2_stat_deadly_mean[M2_stat_deadly_mean$variable == "dry_weight_mg",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
 ylim(0, 10)

M2_weight_plot

M2_R_e_ne_plot <- ggplot() +
  theme_classic() + 
  theme(legend.position="none",
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "", 
       x=expression(),
       y=expression()) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3")) +
  facet_grid(~ timepoint) +
  geom_point(data = M2_stat_deadly_mean[M2_stat_deadly_mean$variable == "mean_R_e_ne",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M2_stat_deadly_mean[M2_stat_deadly_mean$variable == "mean_R_e_ne",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
    ylim(0, 2.5)

M2_R_e_ne_plot

M2_area_encapsulation_plot <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_blank(), legend.position="bottom", 
      plot.title = element_text(size=20), legend.text = element_text(size=15), 
      axis.text.y = element_text(size=15), axis.title = element_text(size=15),
      axis.text.x = element_blank(), axis.ticks.x = element_blank(),
      strip.text = element_text(size=10)) +
  labs(title = "", 
       x=expression(),
       y=expression()) +
  guides(color = guide_legend(override.aes = list(size = 5, shape = c(21,NA,21)))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3"),
                     labels = c("leaves", "", "fish food")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3"),
                    labels = c("leaves", "", "fish food")) +
  facet_grid(~ timepoint) +
  geom_point(data = M2_stat_deadly_mean[M2_stat_deadly_mean$variable == "mean_area_encapsulation",], 
             aes(x=treatment, y=mean, col = treatment, fill = treatment), 
             show.legend = TRUE, size=5, stroke = 1.5,
             alpha = 0.5, shape = 21) +
  geom_errorbar(data = M2_stat_deadly_mean[M2_stat_deadly_mean$variable == "mean_area_encapsulation",],
                aes(x=treatment, col = treatment,
                    y=mean, ymin=mean-sd, ymax=mean+sd), 
                show.legend = FALSE,
                size = 1, alpha = 0.5) +
   ylim(-0.03, 0.09)


M2_area_encapsulation_plot

deadly_endpoints_plot <- ggarrange(M1_weight_plot, M2_weight_plot,
                                   M1_R_e_ne_plot, M2_R_e_ne_plot,
                                   M1_area_encapsulation_plot,
                                   M2_area_encapsulation_plot,
                  labels = c("A", "B", "C", "D", "E", "F"),
                  ncol = 2, nrow = 3)

#png(filename="deadly_endpoints_plot.png", width=10, height=15, units="in", res=300)
deadly_endpoints_plot
#dev.off()
```

# Plot all PUFA
## Preparation

```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean <- filter(microcosm_stat, experiment != "M2")
M2_microcosm_mean <- filter(microcosm_stat, experiment != "M1")
```

Add time point M1
```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean$timepoint <- with(M1_microcosm_mean, 
       ifelse(date_endpoint < "2019-05-03", "0 days",
       ifelse("2019-05-03" <= date_endpoint & date_endpoint <= "2019-05-22", 
              "23 days", "44 days")))
```

Add time point M2
For leaves date is time, when leaves were taken out of stream. Leaves used in experiment at other date/time point:

* 2019-06-04 and 2019-06-18: t2 
* 2019-07-02 and 2019-07-16: t3

```{r, message=FALSE, warning=FALSE}
M2_microcosm_mean$timepoint <- with(M2_microcosm_mean, 
       ifelse(date_endpoint == "2019-07-08", "0 days",
       ifelse(date_endpoint == "2019-07-22", "14 days",
       ifelse(date_endpoint == "2019-06-04", "14 days", 
       ifelse(date_endpoint == "2019-06-18", "14 days",
       ifelse(date_endpoint == "2019-07-12", "14 days",
       ifelse(date_endpoint == "2019-07-19", "14 days",      
       ifelse(date_endpoint == "2019-07-02", "21 days",
       ifelse(date_endpoint == "2019-07-16", "21 days", 
              "21 days")))))))))  
```

Assign start spiders to treatments as for NMDS
```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean$treatment <- with(M1_microcosm_mean, 
                                 ifelse(experiment_ID %in% M1_start_algae, "algae",
                                 ifelse(experiment_ID %in% M1_start_oatmeal,
                                        "oatmeal",
                                 ifelse(experiment_ID %in% M1_start_fishfood,
                                        "fishfood",
                                        treatment))))

M2_microcosm_mean$treatment <- with(M2_microcosm_mean, 
                                 ifelse(experiment_ID %in% M2_start_leaves, "leaves",
                                 ifelse(experiment_ID %in% M2_start_oatmeal,
                                        "oatmeal",
                                 ifelse(experiment_ID %in% M2_start_fishfood,
                                        "fishfood",
                                        treatment))))
```

Get mean and sd for PUFAs per organism, treatment, experiment and time point
```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean <- select(M1_microcosm_mean, -c("date_endpoint", "dry_weight_mg",
                                                "experiment_ID",
                                                "spider_ID_experiment_ID"))
M2_microcosm_mean <- select(M2_microcosm_mean, -c("date_endpoint", "dry_weight_mg",
                                                "experiment_ID",
                                                "spider_ID_experiment_ID"))

M1_microcosm_mean <- reshape2::melt(M1_microcosm_mean, 
                      id.vars = c("experiment", "Organism", "treatment",
                                  "timepoint"))
M2_microcosm_mean <- reshape2::melt(M2_microcosm_mean, 
                      id.vars = c("experiment", "Organism", "treatment",
                                  "timepoint"))

M1_microcosm_mean <- ddply(M1_microcosm_mean, c("experiment", "Organism",
                                              "treatment", "timepoint", "variable"),
                          summarise, mean = mean(value), sd = sd(value))
M2_microcosm_mean <- ddply(M2_microcosm_mean, c("experiment", "Organism", "treatment",
                                              "timepoint", "variable"), 
                          summarise, mean = mean(value), sd = sd(value))
```

Rename
```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean$treatment <- gsub("fishfood", "fish food",
                                     M1_microcosm_mean$treatment)
M1_microcosm_mean$Organism <- gsub("fishfood", "fish food", 
                                    M1_microcosm_mean$Organism)
M1_microcosm_mean$Organism <- gsub("chiro", "chironomids", 
                                    M1_microcosm_mean$Organism)

M2_microcosm_mean$treatment <- gsub("fishfood", "fish food",
                                     M2_microcosm_mean$treatment)
M2_microcosm_mean$Organism <- gsub("fishfood", "fish food", 
                                    M2_microcosm_mean$Organism)
M2_microcosm_mean$Organism <- gsub("chiro", "chironomids", 
                                    M2_microcosm_mean$Organism)
```

Add column with trophic level
```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean$trophic_level <- with(M1_microcosm_mean, ifelse(Organism == "spider", "spider", ifelse(Organism == "chironomids", "chironomids", "basic food source")))

M2_microcosm_mean$trophic_level <- with(M2_microcosm_mean, ifelse(Organism == "spider", "spider", ifelse(Organism == "chironomids", "chironomids", "basic food source")))
```

Reorder factors
```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean$treatment <- factor(M1_microcosm_mean$treatment, 
                                       levels = c("algae", "oatmeal", "fish food"))

M1_microcosm_mean$Organism <- factor(M1_microcosm_mean$Organism, 
                                      levels = c("algae", "oatmeal", "fish food",
                                                 "chironomids",
                                                 "spider"))

M1_microcosm_mean$trophic_level <- factor(M1_microcosm_mean$trophic_level, 
                                         levels = c("basic food source",
                                                    "chironomids", "spider"))

M2_microcosm_mean$treatment <- factor(M2_microcosm_mean$treatment, 
                                       levels = c("leaves", "oatmeal", "fish food"))

M2_microcosm_mean$Organism <- factor(M2_microcosm_mean$Organism, 
                                      levels = c("leaves", "oatmeal", "fish food",
                                                 "chironomids",
                                                 "spider"))

M2_microcosm_mean$trophic_level <- factor(M2_microcosm_mean$trophic_level, 
                                         levels = c("basic food source",
                                                    "chironomids", "spider"))
```

Unit treatment and Organsim column
```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean <- unite(M1_microcosm_mean, treat_orga, treatment, Organism, 
                            remove = FALSE)

M2_microcosm_mean <- unite(M2_microcosm_mean, treat_orga, treatment, Organism, 
                            remove = FALSE)

```

Reorder factors
```{r, message=FALSE, warning=FALSE}
M1_microcosm_mean$treat_orga <- factor(M1_microcosm_mean$treat_orga, 
                                      levels = c("algae_algae",
                                                 "oatmeal_oatmeal", 
                                                 "fish food_fish food",
                                                 
                                                 "algae_chironomids",
                                                 "oatmeal_chironomids",
                                                 "fish food_chironomids",
                                                 
                                                 "algae_spider", 
                                                 "oatmeal_spider",
                                                 "fish food_spider"))

M2_microcosm_mean$treat_orga <- factor(M2_microcosm_mean$treat_orga, 
                                      levels = c("leaves_leaves",
                                                 "oatmeal_oatmeal", 
                                                 "fish food_fish food",
                                                 
                                                 "leaves_chironomids",
                                                 "oatmeal_chironomids",
                                                 "fish food_chironomids",
                                                 
                                                 "leaves_spider", 
                                                 "oatmeal_spider",
                                                 "fish food_spider"))
```

Chose 18:3n-3 (ALA), 18:2n-6c (LIN), 20:4n-6 (ARA) and 20:5n-3 (EPA) 
```{r, message=FALSE, warning=FALSE}
M1_microcosm_export <- filter(M1_microcosm_mean, variable %in% 
                            c("18:3n-3", "18:2n-6c", "20:4n-6", "20:5n-3"))

M2_microcosm_export <- filter(M2_microcosm_mean, variable %in% 
                            c("18:3n-3", "18:2n-6c", "20:4n-6", "20:5n-3"))
```

## Plot
```{r, message=FALSE, warning=FALSE}
M1_LIN_ALA_EPA_ARA <- ifelse(M1_microcosm_mean$variable  %in% 
                            c("18:3n-3", "18:3n-6", 
                              "18:2n-6c", "18:2n-6t", 
                              "20:4n-6", "20:5n-3",
                              "22:6n-3"), "black",
                            "grey")
M2_LIN_ALA_EPA_ARA <- ifelse(M2_microcosm_mean$variable  %in% 
                            c("18:3n-3", "18:3n-6", 
                              "18:2n-6c", "18:2n-6t",
                              "20:4n-6", "20:5n-3",
                              "22:6n-3"), "black",
                            "grey")

M1_all_pufa <- ggplot() +
  theme_classic() + 
  theme(legend.position="bottom",
      plot.title = element_text(size=20), legend.text = element_text(size=20),
      legend.title = element_text(size=20),
      axis.text.y = element_text(size=15, colour = M1_LIN_ALA_EPA_ARA), 
      axis.title = element_text(size=20),
      axis.text.x = element_text(size=20),
      strip.text = element_text(size=20)) +
  labs(title = "Experiment 1", 
       x=expression(~ proportion ~ of ~ total ~ PUFA),
       y=expression()) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkgreen", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkgreen", "blue", "cyan3")) +
 facet_grid(timepoint ~ trophic_level) +
  geom_bar(data = M1_microcosm_mean, 
             aes(x=mean, y=variable, col = treatment, 
             fill = treatment), position = "dodge",
             show.legend = TRUE, alpha = 0.8, lwd=0.1, stat = "identity") +
  geom_errorbar(data = M1_microcosm_mean,
                aes(x=mean, col = treatment,
                    y=variable, xmin=mean, xmax=mean+sd), 
                show.legend = FALSE, alpha = 0.8, position = "dodge")
#png(filename="M1_all_pufa.png", width=15, height=15, units="in", res=300)
M1_all_pufa
# dev.off()

M2_all_pufa <- ggplot() +
  theme_classic() + 
   theme(legend.position="bottom",
      plot.title = element_text(size=20), legend.text = element_text(size=20),
      legend.title = element_text(size=20),
      axis.text.y = element_text(size=15, colour = M2_LIN_ALA_EPA_ARA), 
      axis.title = element_text(size=20),
      axis.text.x = element_text(size=20),
      strip.text = element_text(size=20)) +
  labs(title = "Experiment 2", 
       x=expression(~ proportion ~ of ~ total ~ PUFA),
       y=expression()) +
  guides(color = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values=c("darkolivegreen3", "blue", "cyan3")) +
  scale_fill_manual(values = c("darkolivegreen3", "blue", "cyan3")) +
 facet_grid(timepoint ~ trophic_level) +
  geom_bar(data = M2_microcosm_mean, 
             aes(x=mean, y=variable, col = treatment, 
             fill = treatment), position = "dodge",
             show.legend = TRUE, alpha = 0.8, lwd=0.1, stat = "identity") +
  geom_errorbar(data = M2_microcosm_mean,
                aes(x=mean, col = treatment,
                    y=variable, xmin=mean, xmax=mean+sd), 
                show.legend = FALSE, alpha = 0.8, position = "dodge")

#png(filename="M2_all_pufa.png", width=15, height=15, units="in", res=300)
M2_all_pufa
#dev.off()
```


# Plot temperature variation during experiments
## Preparation

* Temperature in °C

Load data
```{r, message=FALSE, warning=FALSE}
M1_temp_files = list.files(pattern="*_temp_microcosm_1", recursive = TRUE)

M1_temp_data_list = lapply(setNames(M1_temp_files, make.names(gsub("temp_microcosm_1/|_temp_microcosm_1.csv", "", M1_temp_files))), read.csv, stringsAsFactors=TRUE)

M2_temp_files = list.files(pattern="*_temp_microcosm_2.csv", recursive = TRUE)

M2_temp_data_list = lapply(setNames(M2_temp_files, make.names(gsub("temp_microcosm_2/|_temp_microcosm_2.csv", "", M2_temp_files))), read.csv, stringsAsFactors=TRUE)
```

Add name of aerarium to data and get data frame
```{r, message=FALSE, warning=FALSE}
for (i in c(1:16))
{
M1_temp_data_list[[i]]$aerarium <- with(M1_temp_data_list[[i]], names(M1_temp_data_list[i]))
}

M1_temp <- do.call(rbind, M1_temp_data_list)

for (i in c(1:16))
{
M2_temp_data_list[[i]]$aerarium <- with(M2_temp_data_list[[i]], names(M2_temp_data_list[i]))
}

M2_temp <- do.call(rbind, M2_temp_data_list)

M2_temp$aerarium <- gsub("X", "", M2_temp$aerarium)
```

Add experiment
```{r, message=FALSE, warning=FALSE}
M1_temp$experiment <- "M1"

M2_temp$experiment <- "M2"
```

Bind temperature data
```{r, message=FALSE, warning=FALSE}
temp <- rbind(M1_temp, M2_temp)
```

Rename Datum.Zeit
```{r, message=FALSE, warning=FALSE}
names(temp)[2] <- "date_time"
```

Put date_time into POSIXct format
```{r, message=FALSE, warning=FALSE}
temp$date_time <- parse_date_time(temp$date_time, "%m.%d.%y %I:%M:%S %p")
```

Round time on day
```{r, message=FALSE, warning=FALSE}
temp$date_time <- round_date(temp$date_time, unit = "day")
```

Remove NAs in temperature
```{r, message=FALSE, warning=FALSE}
temp <- filter(temp, !is.na(Temp))
```

Get mean and sd per date and time
```{r, message=FALSE, warning=FALSE}
temp_mean <- ddply(temp, c("date_time", "experiment"), 
                   summarise, mean_temp = mean(Temp), sd_temp = sd(Temp))
```

Remove date_time, because experiment ended before
```{r, message=FALSE, warning=FALSE}
#temp_mean <- filter(temp_mean, date_time < "2019-07-29 16:00:00")
temp_mean <- filter(temp_mean, date_time < "2019-07-30")
```

Add day of the experiment
```{r, message=FALSE, warning=FALSE}
temp_mean$day <- with(temp_mean, round(ifelse(experiment == "M1",  difftime(date_time, "2019-04-30", units="days"), difftime(date_time, "2019-07-08", units="days")), digits = 0))
```

Rename experiment
```{r, message=FALSE, warning=FALSE}
temp_mean$experiment <- gsub("M", "", temp_mean$experiment)
```

## Plot temperature profile
```{r, message=FALSE, warning=FALSE}
temp_plot <- ggplot() +
  theme_classic() + 
  theme(legend.title = element_text(size=15), legend.position="bottom", plot.title = element_text(size=20), legend.text = element_text(size=20), axis.text = element_text(size=15), axis.title = element_text(size=15)) +
guides(color = guide_legend(override.aes = list(size = 5))) +
scale_color_manual(values=c("darkgreen", "blue")) +
scale_fill_manual(values = c("darkgreen", "blue")) +
geom_ribbon(data=temp_mean, 
              aes(x=day, y=mean_temp, fill = experiment,
                  ymin=(mean_temp - sd_temp), ymax=(mean_temp + sd_temp)), 
              alpha = 0.1, show.legend = TRUE) +
geom_line(data=temp_mean, 
            aes(x=day, y=mean_temp, col = experiment)) +
  labs(title = "Temperature profile during experiments", 
       x=expression(day ~ of ~ experiment),
       y=expression(temperature ~"("~"°"~ ~C~")")) +
  ylim(0,45)

#png(filename="temp_plot.png", width=10, height=5, units="in", res=300)
temp_plot
#dev.off()
```