---
title: "Trophic transfer of polyunsaturated fatty acids across the aquatic-terrestrial interface: an experimental tritrophic food chain approach: PUFA concentration"
output:
  html_document:
    figure_caption: yes
    highlight: tango
    number_sections: yes
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```
# Required packages
```{r, message=FALSE, warning=FALSE}
library(data.table)
#library(anytime)

library(dplyr) # Data wrangling
library(tidyr) # Data wrangling
library(plyr) # Data wrangling
library(ggplot2) # plots
```

# Set path
```{r, message=FALSE, warning=FALSE}
path <-"/Users/katharinafrisch/Nextcloud/Uni/Paper/Microcosm Paper/Oikos_submission/1_submission/ohler_microcosm/concentration_pufa"
setwd(path)
```
# Calibration
## Prepare data

Load data of standards 
```{r, message=FALSE, warning=FALSE}
standards <- read.table("standards.csv", header = TRUE, sep = ",",na.strings = c("NR", "NA"), stringsAsFactors = FALSE, strip.white = TRUE) 
```
Calibration 5 (04.02.2020): Data for S9-S11 missing, because measurements did not work

Remove failed standard measurements

* 28.01.2020 22_41_1555: batch 3a to check retention time shift, but bad quality

* 06.01.2020 13_14_143: S2 from batch 1, very bad quality 

* 05.02.2020 08_55_338: S8 from batch 5, very bad quality (GC broke during measurement)

* 26.11.2020 19_00_5713: S2 from batch 12, very bad quality


```{r, message=FALSE, warning=FALSE}
standards <- filter(standards, Time != "28.01.2020 22_41_1555")
standards <- filter(standards, Time != "06.01.2020 13_14_143")
standards <- filter(standards, Time != "05.02.2020 08_55_338")
standards <- filter(standards, Time != "26.11.2020 19_00_5713")
```

Check calibration curve from 04.02 - 05.02.2020

* generally carefully‚ with calibration curve of 04.02.2020 - 05.02.2020, since GC broke during this calibration

Check of chromatograms showed:
* "04.02.2020 11_22_591": Bad quality, do not use  
* "04.02.2020 12_18_262": first peaks bad quality  
* "04.02.2020 16_09_493": okay  
* "04.02.2020 17_06_434": first peaks bad quality 
* "04.02.2020 18_02_265": okay  
* "04.02.2020 18_58_066": okay, but first peaks could be better  
* "04.02.2020 19_53_457": okay, but first peaks could be better  

* "05.02.2020 23_45_3727": okay, but no part of calibration curve (can only be used to check RT shift, since GC was broken after S8 measurement)

* to identify PUFA in samples only chromatograms with better quality will be used:
* Remove: 04.02.2020 11_22_591, 04.02.2020 12_18_262, 04.02.2020 17_06_434

```{r, message=FALSE, warning=FALSE}
standards <- filter(standards, Time != "04.02.2020 11_22_591")
standards <- filter(standards, Time != "04.02.2020 12_18_262")
standards <- filter(standards, Time != "04.02.2020 17_06_434")
```

Name peaks in standards
Count how many peaks detected (info needed to split table in list)
```{r, message=FALSE, warning=FALSE}
count <- count(standards, c("Time"))
names(count)[2] <- c("no_peaks")
seq <- c(count$no_peaks)
standards <- left_join(standards, count, by = "Time")

"%notlike%" <- Negate("%like%")

time_06_jan <- filter(standards, Time %like% "06.01.2020")
time_jan_feb_mar <- filter(standards, Time %like% "03.2020" | Time %like% "02.2020" | Time %like% "01.2020")
time_jan_feb_mar <- filter(time_jan_feb_mar, Time %notlike% "06.01.2020")
time_may <- filter(standards, Time %like% "05.2020") 
time_nov <- filter(standards, Time %like% "11.2020") 

time_06_jan <- c(unique(time_06_jan$Time)) # 11
time_jan_feb_mar <- c(unique(time_jan_feb_mar$Time)) # 74
time_may <- c(unique(time_may$Time)) # 15
time_nov <- c(unique(time_nov$Time)) # 26
time_may_nov <- c(time_may, time_nov)

length((unique(standards$Time))) #126: sum of single times add up to sum of all times
```

Get list
```{r, message=FALSE, warning=FALSE}
standard_list <- split(standards, f = c(standards$Time), seq(along.with = seq))
```

Sort by Time and decreasing RT
```{r, message=FALSE, warning=FALSE}
for (i in c(1:126))
{
standard_list[[i]] <- arrange(standard_list[[i]], desc(RT))
}
```

Get name of PUFAs in right order
```{r, message=FALSE, warning=FALSE}
name_34 <- c("24:1n-9", "24:0", "22:6n-3", "23:0", "22:2n-6", "22:1n-9", "22:0", "20:5n-3", "20:3n-3", "21:0", "20:4n-6", "20:3n-6", "20:2n-6", "20:1n-9", "20:0", "18:3n-3", "18:3n-6", "18:2n-6c", "18:2n-6t", "18:1n-7", "18:1n-9c", "18:1n-9t", "18:0", "17:1n-7", "17:0", "16:1n-7", "16:0", "15:1n-5", "15:0", "14:1n-5", "14:0", "13:0", "12:0", "11:0")

name_33 <- c("24:1n-9", "24:0", "22:6n-3", "23:0", "22:2n-6", "22:1n-9", "22:0", "20:5n-3", "20:3n-3", "21:0", "20:4n-6", "20:3n-6", "20:2n-6", "20:1n-9", "20:0", "18:3n-3", "18:3n-6", "18:2n-6c", "18:2n-6t", "18:1n-7", "18:1n-9c", "18:1n-9t", "18:0", "17:1n-7", "17:0", "16:1n-7", "16:0", "15:1n-5", "15:0", "14:1n-5", "14:0", "13:0", "12:0")

name_27 <- c("24:1n-9", "24:0", "22:6n-3", "23:0", "22:2n-6", "22:1n-9", "22:0", "20:5n-3", "20:3n-3", "21:0", "20:4n-6", "20:3n-6", "20:2n-6", "20:1n-9", "20:0", "18:3n-3", "18:3n-6", "18:2n-6c", "18:2n-6t", "18:1n-7", "18:1n-9c", "18:1n-9t", "18:0", "17:1n-7", "17:0", "16:1n-7", "16:0") # measurements in November (GC was broken inbetween and only PUFA higher than 16:0 could be identified)

name_26 <- c("24:1n-9", "24:0", "22:6n-3", "23:0", "22:2n-6", "22:1n-9", "22:0", "20:5n-3", "20:3n-3", "21:0", "20:4n-6", "20:3n-6", "20:2n-6", "20:1n-9", "20:0", "18:3n-3", "18:3n-6", "18:2n-6c", "18:2n-6t", "18:1n-7", "18:1n-9c", "18:1n-9t", "18:0", "17:1n-7", "17:0", "16:1n-7") # measurement S1 batch 12 in November 
```

Add column with PUFA names
```{r, message=FALSE, warning=FALSE}
for (i in c(1:126))
{
standard_list[[i]]$name_pufa <- with(standard_list[[i]], 
                                     ifelse(no_peaks == 34, name_34, 
                                     ifelse(no_peaks == 33, name_33, 
                                     ifelse(no_peaks == 27, name_27,
                                     ifelse(no_peaks == 26, name_26,      
                                            "NA")))))
}
```

Add c1, V1 and V2 (dilution for calibration)

* Supelco contains 37 FAMEs, we start quantification at C11:00 for measurements before November 2020 and after November 2020 at C16:0

* 18:1n-7 ((11E)-11-octadecenoic acid methyl acid) manually added, because it occurs in samples and is not part of Supelco

* 18:3n-3: manually added, it already is in Supelco, but concentration in Supelco lower as in samples 

* 18:3n-3: in Supelco 201 microgramm per mL and manually added 223.75 microgramm per mL
-> thus put at first "1" in column c1 as "place holder" and then calculation manually (in libre office, otherwise too much ifelse and so on)

Add c1 (mass concentration of FAME in undiluted Supelco/standard) in microgramm per mL
```{r, message=FALSE, warning=FALSE}
for (i in c(1:126))
{
standard_list[[i]]$c1 <- with(standard_list[[i]], 
                              ifelse(name_pufa == "11:0", 200, 
                              ifelse(name_pufa == "12:0", 401, 
                              ifelse(name_pufa == "13:0", 201, 
                              ifelse(name_pufa == "14:0", 401, 
                              ifelse(name_pufa == "14:1n-5", 201, 
                              ifelse(name_pufa == "15:0", 200, 
                              ifelse(name_pufa == "15:1n-5", 200, 
                              ifelse(name_pufa == "16:0", 601, 
                              ifelse(name_pufa == "16:1n-7", 201, 
                              ifelse(name_pufa == "17:0", 201,
                              ifelse(name_pufa == "17:1n-7", 200, 
                              ifelse(name_pufa == "18:0", 401, 
                              ifelse(name_pufa == "18:1n-9t", 200,
                              ifelse(name_pufa == "18:1n-9c", 401,
                              ifelse(name_pufa == "18:1n-7", 200,
                              ifelse(name_pufa == "18:2n-6t", 201, 
                              ifelse(name_pufa == "18:2n-6c", 201,
                              ifelse(name_pufa == "18:3n-6", 201,
                              ifelse(name_pufa == "18:3n-3", 1,
                              ifelse(name_pufa == "20:0", 401, 
                              ifelse(name_pufa == "20:1n-9", 201, 
                              ifelse(name_pufa == "20:2n-6", 200, 
                              ifelse(name_pufa == "20:3n-6", 202, 
                              ifelse(name_pufa == "20:4n-6", 201, 
                              ifelse(name_pufa == "21:0", 200, 
                              ifelse(name_pufa == "20:3n-3", 200, 
                              ifelse(name_pufa == "20:5n-3", 201, 
                              ifelse(name_pufa == "22:0", 401, 
                              ifelse(name_pufa == "22:1n-9", 201, 
                              ifelse(name_pufa == "22:2n-6", 200,  
                              ifelse(name_pufa == "23:0", 200, 
                              ifelse(name_pufa == "22:6n-3", 201, 
                              ifelse(name_pufa == "24:0", 401, 
                              ifelse(name_pufa == "24:1n-9", 200,
                              NA)))))))))))))))))))))))))))))))))))
}
```

Add V1 (volume Supelco/standard used for dilution) in mL
```{r, message=FALSE, warning=FALSE}
for (i in c(1:126))
{
standard_list[[i]]$V1 <- with(standard_list[[i]],
                         ifelse(Time %like% "06.01.2020" & 
                                  runinfo %like% "S1 Supelco", 0.00125, 
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S2 Supelco", 0.00250,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S3 Supelco", 0.00125,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S4 Supelco", 0.00250,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S5 Supelco", 0.00375,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S6 Supelco", 0.00500,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S7 Supelco", 0.00625,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S8 Supelco", 0.00750,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S9 Supelco", 0.00875,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S10 Supelco", 0.01000,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S11 Supelco", 0.01130,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S12 Supelco", 0.01250,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S1 Supelco", 0.00500, 
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S2 Supelco", 0.00500,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S3 Supelco", 0.00750,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S4 Supelco", 0.00500,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S5 Supelco", 0.00630,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S6 Supelco", 0.00750,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S7 Supelco", 0.00880,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S8 Supelco", 0.01000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S9 Supelco", 0.01130,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S10 Supelco", 0.01250,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S11 Supelco", 0.01380,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S12 Supelco", 0.01500,
                         ifelse(Time == "25.11.2020 02_15_0314" &
                                  runinfo %like% "S13 Supelco", 0.0188,   
                         ifelse(Time == "27.11.2020 04_31_1224" & 
                                  runinfo %like% "S13 Supelco", 0.015,        
                         ifelse(runinfo == "S6 A Supelco and 2 FAME", 0.005, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S1 Supelco", 0.0008,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S2 Supelco", 0.0015,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S3 Supelco", 0.0023,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S4 Supelco", 0.0030,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S5 Supelco", 0.0038,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S6 Supelco", 0.0045,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S7 Supelco", 0.0053,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S8 Supelco", 0.0060,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S9 Supelco", 0.0068,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S10 Supelco", 0.0075, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S11 Supelco", 0.0083, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S12 Supelco", 0.009, 
                                NA))))))))))))))))))))))))))))))))))))))))
}                                            
```

Add V2 (end volume of dilution/standard) in mL
```{r, message=FALSE, warning=FALSE}
for (i in c(1:126))
{
standard_list[[i]]$V2 <- with(standard_list[[i]],
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S1 Supelco", 0.05005, 
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S2 Supelco", 0.05000,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S3 Supelco", 0.05005,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S4 Supelco", 0.05000,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S5 Supelco", 0.05005,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S6 Supelco", 0.05000,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S7 Supelco", 0.05005,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S8 Supelco", 0.05000,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S9 Supelco", 0.05005,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S10 Supelco", 0.05000,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S11 Supelco", 0.05020,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S12 Supelco", 0.05000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S1 Supelco", 0.20000, 
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S2 Supelco", 0.10000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S3 Supelco", 0.10000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S4 Supelco", 0.05000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S5 Supelco", 0.05040,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S6 Supelco", 0.05000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S7 Supelco", 0.05020,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S8 Supelco", 0.05000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S9 Supelco", 0.05020,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S10 Supelco", 0.05000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S11 Supelco", 0.05020,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S12 Supelco", 0.05000,
                         ifelse(Time == "25.11.2020 02_15_0314" &
                                  runinfo %like% "S13 Supelco", 0.0764,   
                         ifelse(Time == "27.11.2020 04_31_1224" & 
                                  runinfo %like% "S13 Supelco", 0.045,  
                         ifelse(runinfo == "S6 A Supelco and 2 FAME", 0.032, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S1 Supelco", 0.03020,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S2 Supelco", 0.03000,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S3 Supelco", 0.03020,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S4 Supelco", 0.03000,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S5 Supelco", 0.03020,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S6 Supelco", 0.03000,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S7 Supelco", 0.03020,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S8 Supelco", 0.03000,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S9 Supelco", 0.03020,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S10 Supelco", 0.03000, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S11 Supelco", 0.03020, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S12 Supelco", 0.03000,
                                NA))))))))))))))))))))))))))))))))))))))))
}    
```

Add c2 (concentration in diluted supelco/standard) microgramm per mL

For 18:3n-3 c2 calculated in libre office and then added (otherwise too many ifelse), for other FAME calculated here in R:

$$ c_2 = {c_1 \cdot V_1 \over V_2} $$

```{r, message=FALSE, warning=FALSE}
for (i in c(1:126))
{
standard_list[[i]]$c2 <- with(standard_list[[i]],
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S1 Supelco" & 
                                  name_pufa == "18:3n-3", 1.06081, 
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S2 Supelco" &
                                  name_pufa == "18:3n-3", 2.12375,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S3 Supelco" &
                                  name_pufa == "18:3n-3", 10.60814,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S4 Supelco" &
                                  name_pufa == "18:3n-3", 21.23750,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S5 Supelco" &
                                  name_pufa == "18:3n-3", 31.82443,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S6 Supelco" &
                                  name_pufa == "18:3n-3", 42.47500,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S7 Supelco" &
                                  name_pufa == "18:3n-3", 53.04071,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S8 Supelco" &
                                  name_pufa == "18:3n-3", 63.71250,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S9 Supelco" &
                                  name_pufa == "18:3n-3", 74.25699,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S10 Supelco" &
                                  name_pufa == "18:3n-3", 84.95000,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S11 Supelco" &
                                  name_pufa == "18:3n-3", 95.61106,
                         ifelse(Time %like% "06.01.2020" &
                                  runinfo %like% "S12 Supelco" &
                                  name_pufa == "18:3n-3", 106.18750,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S1 Supelco" &
                                  name_pufa == "18:3n-3", 10.61875, 
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S2 Supelco" &
                                  name_pufa == "18:3n-3", 21.23750,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S3 Supelco" &
                                  name_pufa == "18:3n-3", 31.85625,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S4 Supelco" &
                                  name_pufa == "18:3n-3", 42.47500,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S5 Supelco" &
                                  name_pufa == "18:3n-3", 53.09375,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S6 Supelco" &
                                  name_pufa == "18:3n-3", 63.71250,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S7 Supelco" &
                                  name_pufa == "18:3n-3", 74.45817,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S8 Supelco" &
                                  name_pufa == "18:3n-3", 84.95000,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S9 Supelco" &
                                  name_pufa == "18:3n-3", 95.61106,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S10 Supelco" &
                                  name_pufa == "18:3n-3", 106.18750,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S11 Supelco" &
                                  name_pufa == "18:3n-3", 116.76394,
                         ifelse(Time %in% time_may_nov &
                                  runinfo %like% "S12 Supelco" &
                                  name_pufa == "18:3n-3", 127.42500,
                         ifelse(Time == "25.11.2020 02_15_0314" &
                                  runinfo %like% "S13 Supelco", 104.51963,   
                         ifelse(Time == "27.11.2020 04_31_1224" & 
                                  runinfo %like% "S13 Supelco", 141.58333,
                         ifelse(runinfo == "S6 A Supelco and 2 FAME" &
                                  name_pufa == "18:3n-3", 45.18617, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S1 Supelco" &
                                  name_pufa == "18:3n-3", 11.25166,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S2 Supelco" &
                                  name_pufa == "18:3n-3", 21.23750,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S3 Supelco" &
                                  name_pufa == "18:3n-3", 32.34851,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S4 Supelco" &
                                  name_pufa == "18:3n-3", 42.47500,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S5 Supelco" &
                                  name_pufa == "18:3n-3", 53.44536,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S6 Supelco" &
                                  name_pufa == "18:3n-3", 63.71250,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S7 Supelco" &
                                  name_pufa == "18:3n-3", 74.54222,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S8 Supelco" &
                                  name_pufa == "18:3n-3", 84.95000,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S9 Supelco" &
                                  name_pufa == "18:3n-3", 95.63907,
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S10 Supelco" &
                                  name_pufa == "18:3n-3", 106.18750, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S11 Supelco" &
                                  name_pufa == "18:3n-3", 116.73593, 
                         ifelse(Time %in% time_jan_feb_mar &
                                  runinfo %like% "S12 Supelco" &
                                  name_pufa == "18:3n-3", 127.42500,
                         (c1*V1)/V2))))))))))))))))))))))))))))))))))))))))
}  
```

For calibration curve 06.01.2020 12_17_082 we need to put concentration of S6 as c1 and calculate c2 with right value (c1 was taken from S6 and not undiluted standards)
```{r, message=FALSE, warning=FALSE}
standard_list[["06.01.2020 12_17_082"]]$c1 <- standard_list[["06.01.2020 17_52_418"]]$c2

standard_list[["06.01.2020 12_17_082"]]$c2 <- with(standard_list[["06.01.2020 12_17_082"]], (c1*V1)/V2)
```

## Calculate calibration curves (linear)

$$ area = a \cdot {concentration } + b $$

* Get slope (a), intercept (b) and R^2

Get data frame from list
```{r, message=FALSE, warning=FALSE}
standards <- do.call(rbind, standard_list) 
```

Get standards for Jan - Mar 2020, May - Jun 2020, Nov 2020
```{r, message=FALSE, warning=FALSE}
standards$cali <-with(standards, ifelse(Date > "2020-01-01" & Date <= "2020-04-01","cali_1",ifelse(Date > "2020-05-01" & Date <= "2020-07-01","cali_2","cali_3")))
```

Get info needed to get lists with single calibration curves
Need info of date and name_pufa
```{r, message=FALSE, warning=FALSE}
standards <- unite(standards, "cali_pufa", "cali", "name_pufa", sep = "_", remove = FALSE)

count_cali <- count(standards, c("cali_pufa"))
names(count_cali)[2] <- c("no_measurements_cali")
seq <- c(count_cali$no_measurements_cali)
calibration <- left_join(standards, count_cali, by = "cali_pufa")

#unique(calibration$Date)
#unique(calibration$cali_pufa)
```

Get list
```{r, message=FALSE, warning=FALSE}
calibration_list <- split(calibration, f = c(calibration$cali_pufa), seq(along.with = seq))
```

Get slope
```{r, message=FALSE, warning=FALSE}
for (i in c(1:94)) 
 {
calibration_list[[i]]$slope <- with(calibration_list[[i]], ifelse(nrow(calibration_list[[i]]) > 1, lm(area ~ c2, data = calibration_list[[i]])[[1]][[2]], NA))
}
```

Get intercept
```{r, message=FALSE, warning=FALSE}
for (i in c(1:94)) 
 {
calibration_list[[i]]$intercept <- with(calibration_list[[i]], ifelse(nrow(calibration_list[[i]]) > 1, lm(area ~ c2, data = calibration_list[[i]])[[1]][[1]], NA))
}
```

Get adjusted r squared
```{r, message=FALSE, warning=FALSE}
for (i in c(1:94)) 
 {
calibration_list[[i]]$adj.r.squared <- with(calibration_list[[i]], ifelse(nrow(calibration_list[[i]]) > 1, summary(lm(area ~ c2, data = calibration_list[[i]]))$adj.r.squared, NA))
}
```

### Check calibrations

Check $R^2$ < 0.7 for PUFA with 17 or more C-Atoms (other PUFA not used in analysis)

Get PUFAs
```{r, message=FALSE, warning=FALSE}
calibration <- do.call(rbind, calibration_list)
name_17 <- c("24:1n-9", "24:0", "22:6n-3", "23:0", "22:2n-6", "22:1n-9", "22:0", "20:5n-3", "20:3n-3", "21:0", "20:4n-6", "20:3n-6", "20:2n-6", "20:1n-9", "20:0", "18:3n-3", "18:3n-6", "18:2n-6c", "18:2n-6t", "18:1n-7", "18:1n-9c", "18:1n-9t", "18:0", "17:1n-7", "17:0")

check_r2 <- filter(calibration, name_pufa %in% name_17)
check_r2 <- select(check_r2, c("cali_pufa", "name_pufa", "adj.r.squared"))
```

Get R^2 for calibrations
```{r, message=FALSE, warning=FALSE}
check_r2 <- distinct(check_r2)
check_r2$adj.r.squared <- round(check_r2$adj.r.squared, digits = 1)

check_r2_0.7 <- filter(check_r2, adj.r.squared < 0.7)

name_check_r2_0.7 <- unique(check_r2_0.7$cali_pufa)
```

Visual check for calibrations with R^2 < 0.7
Get data
```{r, message=FALSE, warning=FALSE}
check_cali_r2_0.7 <- filter(calibration, cali_pufa %in% name_check_r2_0.7)

name_check_cali_1_r2_0.7 <- c("cali_1_17:0", "cali_1_17:1n-7", "cali_1_18:1n-9c", "cali_1_18:2n-6t", "cali_1_20:1n-9", "cali_1_20:3n-3", "cali_1_24:0") 

name_check_cali_3_r2_0.7 <- c("cali_3_17:0", "cali_3_17:1n-7", "cali_3_18:1n-7", "cali_3_18:1n-9c", "cali_3_18:1n-9t", "cali_3_18:2n-6c", "cali_3_18:2n-6t", "cali_3_18:3n-3", "cali_3_18:3n-6",  "cali_3_20:0", "cali_3_20:1n-9", "cali_3_20:2n-6", "cali_3_20:3n-3", "cali_3_20:3n-6", "cali_3_20:4n-6", "cali_3_20:5n-3", "cali_3_21:0", "cali_3_22:0", "cali_3_22:1n-9", "cali_3_22:2n-6", "cali_3_22:6n-3", "cali_3_23:0", "cali_3_24:0", "cali_3_24:1n-9" )
```

Plot Calibration 1
```{r, message=FALSE, warning=FALSE}
for (i in name_check_cali_1_r2_0.7)
 {
plot(calibration_list[[i]]$area ~ calibration_list[[i]]$c2, yaxt="n", ylab="", pch = 16, xlab ="", main = i)
axis(2, las=2)
title(ylab= expression(Integrated~area), xlab = expression(concentration~microgramm~mL))
abline(lm(calibration_list[[i]]$area ~ calibration_list[[i]]$c2), lwd = 2)
text(60, 100, calibration_list[[i]]$adj.r.squared[1])
}
```

For cali 1 and 2: Force calibration through 0 to improve R^2

Plot calibration 3
```{r, message=FALSE, warning=FALSE}
for (i in name_check_cali_3_r2_0.7)
 {
plot(calibration_list[[i]]$area ~ calibration_list[[i]]$c2, yaxt="n", ylab="", pch = 16, xlab ="", main = i)
axis(2, las=2)
title(ylab= expression(Integrated~area), xlab = expression(concentration~microgramm~mL))
abline(lm(calibration_list[[i]]$area ~ calibration_list[[i]]$c2), lwd = 2)
text(120, 150000, calibration_list[[i]]$adj.r.squared[1])
}
```

Force calibration through 0, remove two highest concentrations and outlier around 40 microgramm/mL

## Calculate calibration curves to improve R^2

$$ area = a \cdot {concentration } + 0 $$

* Get slope (a) and R^2

Get data frame from list
```{r, message=FALSE, warning=FALSE}
standards_2 <- do.call(rbind, standard_list) 
```

Get standards for Jan - Mar 2020, May - Jun 2020, Nov 2020
```{r, message=FALSE, warning=FALSE}
standards_2$cali <-with(standards_2, ifelse(Date > "2020-01-01" & Date <= "2020-04-01","cali_1",ifelse(Date > "2020-05-01" & Date <= "2020-07-01","cali_2","cali_3")))
```

Get info needed to get lists with single calibration curves
Need info of date and name_pufa
```{r, message=FALSE, warning=FALSE}
standards_2 <- unite(standards_2, "cali_pufa", "cali", "name_pufa", sep = "_", remove = FALSE)

count_cali <- count(standards_2, c("cali_pufa"))
names(count_cali)[2] <- c("no_measurements_cali")
seq <- c(count_cali$no_measurements_cali)
calibration_2 <- left_join(standards_2, count_cali, by = "cali_pufa")

#unique(calibration_2$Date)
#unique(calibration_2$cali_pufa)
```

Get list
```{r, message=FALSE, warning=FALSE}
calibration_list_2 <- split(calibration_2, f = c(calibration_2$cali_pufa), seq(along.with = seq))
```

Remove two highest concentrations of cali_3
```{r, message=FALSE, warning=FALSE}
remove <- filter(calibration_2, cali == "cali_3")
remove <- unique(remove$cali_pufa)

for (i in remove) 
 {
calibration_list_2[[i]] <- filter(calibration_list_2[[i]], c2 < max(c2))
}

for (i in remove) 
 {
calibration_list_2[[i]] <- filter(calibration_list_2[[i]], c2 < max(c2))
}
```

Remove single measurement with high concentration around 40 microgram/mL
```{r, message=FALSE, warning=FALSE}
for (i in c(1:94)) 
 {
calibration_list_2[[i]] <- filter(calibration_list_2[[i]], 
                                  runname != "AcqMethodName24.11.2020 12_09_562")
}
```

Get slope and force intercept through 0
```{r, message=FALSE, warning=FALSE}
for (i in c(1:94)) 
 {
calibration_list_2[[i]]$slope <- with(calibration_list_2[[i]], ifelse(nrow(calibration_list_2[[i]]) > 1, lm(area ~ 0 + c2, data = calibration_list_2[[i]])[[1]][[1]], NA))
}
```

Get adjusted r squared
```{r, message=FALSE, warning=FALSE}
for (i in c(1:94)) 
 {
calibration_list_2[[i]]$adj.r.squared <- with(calibration_list_2[[i]], ifelse(nrow(calibration_list_2[[i]]) > 1, summary(lm(area ~ c2, data = calibration_list_2[[i]]))$adj.r.squared, NA))
}
```

### Check calibrations

Check R^2 for PUFA with 17 or more C-Atoms
Get PUFAs
```{r, message=FALSE, warning=FALSE}
calibration_2 <- do.call(rbind, calibration_list_2)

check_r2_2 <- filter(calibration_2, name_pufa %in% name_17)
check_r2_2 <- select(check_r2_2, c("cali_pufa", "name_pufa", "adj.r.squared"))
```

Get R^2 for calibrations
```{r, message=FALSE, warning=FALSE}
check_r2_2 <- distinct(check_r2_2)
check_r2_2$adj.r.squared <- round(check_r2_2$adj.r.squared, digits = 1)
names(check_r2_2)[3] <- "adj.r.squared_2"
```

Compare R^2
Merge $R^2$ data
```{r, message=FALSE, warning=FALSE}
check_r2_2 <- left_join(check_r2, check_r2_2, by =c("cali_pufa", "name_pufa"))
```

Check if $R^2$ is higher after improvement measures
```{r, message=FALSE, warning=FALSE}
filter(check_r2_2, adj.r.squared > adj.r.squared_2)

filter(check_r2_2, adj.r.squared < adj.r.squared_2) 
```
Improved for 23 calibrations

Visual check for calibrations 

-> plot calibration original and changed in same plot with different colours to compare
-> blue original calibration
-> black calibration forced through 0 and without outliers

Plot
```{r, message=FALSE, warning=FALSE}
names <- unique(calibration$cali_pufa)

for (i in names)
 {
plot(calibration_list[[i]]$area ~ calibration_list[[i]]$c2, yaxt="n", ylab="", xlab ="", main = i, pch = 0, col = "blue")
axis(2, las=2)
points(x = calibration_list_2[[i]]$c2, y = calibration_list_2[[i]]$area, pch = 16)
title(ylab= expression(Integrated~area), xlab = expression(concentration~microgramm~mL))
abline(lm(calibration_list_2[[i]]$area ~ 0 + calibration_list_2[[i]]$c2), lwd = 2)
abline(lm(calibration_list[[i]]$area ~ calibration_list[[i]]$c2), lwd = 2, col = "blue")
}
```

Get data frame
```{r, message=FALSE, warning=FALSE}
calibration <- do.call(rbind, calibration_list_2)
```

# LOQ
## Prepare data

Get data of blinds
```{r, message=FALSE, warning=FALSE}
blind <- read.table("blind.csv", header = TRUE, sep = ",",na.strings = c("NR", "NA"), stringsAsFactors = FALSE, strip.white = TRUE) 
```

Remove not identified peaks in blinds and blinds to which internal standard was added accidentally 
```{r, message=FALSE, warning=FALSE}
blind <- filter(blind, name_pufa !="")
blind <- filter(blind, Internal_Standard_Volume_ID == 0)
```

Round c2 on one digit
```{r, message=FALSE, warning=FALSE}
standards$c2 <- round(standards$c2, digits = 0)
```

Get LOQ for Jan-Mar, May-Jun, Nov
```{r, message=FALSE, warning=FALSE}
blind$cali <-with(blind, ifelse(Date > "2020-01-01" & Date <= "2020-04-01","cali_1",ifelse(Date > "2020-05-01" & Date <= "2020-07-01","cali_2","cali_3")))

blind <- unite(blind, "cali_pufa", "cali", "name_pufa", sep = "_", remove = FALSE)
```

Get mean area per PUFA and calibration in blinds and standards
```{r, message=FALSE, warning=FALSE}
blind_mean <- ddply(blind, c("cali_pufa"), summarise, mean_area_blind = mean(area))

standards_mean <- ddply(standards, c("cali_pufa", "c2"), summarise, mean_area_standard = mean(area))
```

Merge data
```{r, message=FALSE, warning=FALSE}
LOQ <- merge(standards_mean, blind_mean, by ="cali_pufa")
```


## Calculate LOQ

Check if factor > 3 is between mean blind area and standard

First divide mean area of standards with mean area of blind

$$ LOQ = { \frac {mean\ \ area\ \ standard} {mean\ \ area\ \ blind}} $$

```{r, message=FALSE, warning=FALSE}
LOQ$f <- with(LOQ, round(mean_area_standard/mean_area_blind, digits = 0))
```

Filter to get all factors > 3
```{r, message=FALSE, warning=FALSE}
LOQ <- filter(LOQ, f > 3)
```

Chose per PUFA smallest c2
```{r, message=FALSE, warning=FALSE}
LOQ <- ddply(LOQ, c("cali_pufa"), summarise, min_c2 = min(c2))
```

Some PUFA not found in blinds so, min_c2 = 1/2 concentration of lowest standard.
```{r, message=FALSE, warning=FALSE}
half_conc <- ddply(standards_mean, c("cali_pufa"), summarise, min_c2 = min(c2))

half_conc$min_c2 <- half_conc$min_c2/2

not_zero <- unique(LOQ$cali_pufa)

`%notin%` <- Negate(`%in%`)

half_conc <- filter(half_conc, cali_pufa %notin% not_zero)

LOQ <- rbind(LOQ, half_conc)
```

# Internal standards
## Prepare data

Get data of internal standards
```{r, message=FALSE, warning=FALSE}
si <- read.table("internal_standard.csv", header = TRUE, sep = ",",na.strings = c("NR", "NA"), stringsAsFactors = FALSE, strip.white = TRUE) 
```

mg in microgram
```{r, message=FALSE, warning=FALSE}
si$weighed_in_microg <- with(si, weighed_in_mg*1000)
```

Standard_ID in same format as in checklists
```{r, message=FALSE, warning=FALSE}
si$Standard_ID <- gsub("_", "", si$Standard_ID)
```

Add volume of internal standard added to samples
* for all standards: 50 - 400 microL -> 0.05 mL - 0.4 mL
* Name in columns will be microL, because this is needed to add it to samples
* Value in columns will be mL, because mL used for all calculations
```{r, message=FALSE, warning=FALSE}
si$'50' <- 0.050

si$'100' <- 0.100
si$'150' <- 0.150

si$'200' <- 0.200
si$'250' <- 0.250

si$'300' <- 0.300
si$'350' <- 0.350

si$'400' <- 0.400

si <- reshape2::melt(si, id.vars = c("Lipid", "Standard_ID", "date", "weighed_in_mg" , "Volume_MeOH_mL", "comment", "weighed_in_microg"), variable.name = "volume_added_microL", value.name = "volume_added_mL")
```

Add column Internal_Standard_Volume_ID, which is needed to merge later with samples
50_SI6
```{r, message=FALSE, warning=FALSE}
si <- unite(si, Internal_Standard_Volume_ID, volume_added_microL, 
            Standard_ID, sep = "_")
```

Rename lipid into name_pufa and put in same format as in samples
```{r, message=FALSE, warning=FALSE}
names(si)[1] <- "name_pufa"
si$name_pufa <- gsub("C", "", si$name_pufa)
```

## Calculate c1 of internal standards 

* $c_1$: concentration in microgram per mL after internal standards were dissolved in MeOH
* m: mass of internal standard weighed in 
* $V_1$: Volume of MeOH used to dissolve internal standards

$$ c_1 = {\frac {m} {V_1}}$$
```{r, message=FALSE, warning=FALSE}
si$c1 <- with(si, weighed_in_microg/Volume_MeOH_mL)
```

## Calculate mass of internal standards added to samples

* $V_2$: Volume internal standard added to sample
* m: mass of internal standards added to samples
* $c_1$:concentration in microgram per mL after internal standards were dissolved in MeOH

$$ m = c_1 \cdot V_2 $$

```{r, message=FALSE, warning=FALSE}
si$m_si_in_sample <- with(si, c1*volume_added_mL)
```

Chose columns needed to calculate recovery
```{r, message=FALSE, warning=FALSE}
recovery <- select(si, "name_pufa", "Internal_Standard_Volume_ID", "m_si_in_sample"  )
```

# Recovery

Get potential abbreviations from original concentration of PUFA in samples e.g. losses during extraction and derivatization.

## Prepare data

Get data of samples
```{r, message=FALSE, warning=FALSE}
microcosm <- read.table("microcosm.csv", header = TRUE, sep = ",",na.strings = c("NR", "NA"), stringsAsFactors = FALSE, strip.white = TRUE) 
microcosm_complete <- read.table("microcosm.csv", header = TRUE, sep = ",",na.strings = c("NR", "NA"), stringsAsFactors = FALSE, strip.white = TRUE) 

emergence <- read.table("emergence.csv", header = TRUE, sep = ",",na.strings = c("NR", "NA"), stringsAsFactors = FALSE, strip.white = TRUE) 

spiders <- read.table("spiders_field.csv", header = TRUE, sep = ",",na.strings = c("NR", "NA"), stringsAsFactors = FALSE, strip.white = TRUE) 
```

Rename columns containing dry weight
```{r, message=FALSE, warning=FALSE}
names(spiders)[7] <- "dry_weight_mg"

names(emergence)[4] <- "dry_weight_mg" 
```

Get needed columns and bind experimental data
```{r, message=FALSE, warning=FALSE}
microcosm <- select(microcosm, c("experiment", "GC_ID", "Internal_Standard_Volume_ID", "MeOH_Volume_dissolved", "TMSH_Volume", "Sample_Volume_derivatized", "runinfo", "area", "name_pufa", "runname", "Date", "dry_weight_mg"))                       

spiders <- select(spiders, c("experiment", "GC_ID", "Internal_Standard_Volume_ID", "MeOH_Volume_dissolved", "TMSH_Volume", "Sample_Volume_derivatized", "runinfo", "area", "name_pufa", "runname", "Date", "dry_weight_mg"))  

emergence <- select(emergence, c("experiment", "GC_ID", "Internal_Standard_Volume_ID", "MeOH_Volume_dissolved", "TMSH_Volume", "Sample_Volume_derivatized", "runinfo", "area", "name_pufa", "runname", "Date", "dry_weight_mg"))  

rec_data <- rbind(microcosm, spiders, emergence)
```

Get C23 and C17 data
```{r, message=FALSE, warning=FALSE}
rec_data <- filter(rec_data, name_pufa %in% c("17:0", "23:0"))
```

Get column "cali_name" needed to merge with calibration data 
```{r, message=FALSE, warning=FALSE}
rec_data$cali <-with(rec_data, ifelse(Date > "2020-01-01" & Date <= "2020-04-01","cali_1",ifelse(Date > "2020-05-01" & Date <= "2020-07-01","cali_2","cali_3")))
```

Column with info needed to merge correctly
```{r, message=FALSE, warning=FALSE}
rec_data <- unite(rec_data, "cali_pufa", "cali", "name_pufa", sep = "_", remove = FALSE)
```

Columns needed from standards to calculate conc. in samples
```{r, message=FALSE, warning=FALSE}
cali <- select(calibration, c("cali_pufa", "slope")) 
```

For every standard measurement in same calibration curve and substance slope, which are the same
```{r, message=FALSE, warning=FALSE}
cali <- distinct(cali) 
```

Merge with calibration 
```{r, message=FALSE, warning=FALSE}
rec_data <- left_join(rec_data, cali, by = "cali_pufa")
```

Merge with LOQ table
```{r, message=FALSE, warning=FALSE}
rec_data <- left_join(rec_data, LOQ, by = "cali_pufa")
```

## Calculate 17:0 and 23:0 concentration in samples

PUFA dissolved: Measured concentration in microgramm per mL

$$ area = slope \cdot concentration $$
$$ concentration = {\frac  {area } {slope}} $$

```{r, message=FALSE, warning=FALSE}
rec_data$conc_measured <- with(rec_data, (area)/slope)
```

Include LOQ
Set 0 for concentration < LOQ 
```{r, message=FALSE, warning=FALSE}
rec_data$conc_measured_LOQ <- with(rec_data, ifelse(conc_measured < min_c2, 0, conc_measured))
```

Merge microcosm and recovery data
```{r, message=FALSE, warning=FALSE}
recovery <- left_join(rec_data, recovery, by = c("name_pufa", "Internal_Standard_Volume_ID"))
```

Concentration internal standard in sample dissolved in MeOH after evaporation of Chloroform:MeOH

* $V_2$: Volume MeOH, in which solved (MeOH_Volume_dissolved -> is in microlitre, so divide by 1000 to get mL)
* $c_2$: Concentration internal standard after sample dissolved -> c_si_dissolved -> microgram per mL

$$ c_2 = {\frac{m} {V_2}} $$
```{r, message=FALSE, warning=FALSE}
recovery$c_si_dissolved <- with(recovery, m_si_in_sample/(MeOH_Volume_dissolved/1000))
```

Get measured concentration of internal standard
* $V_3$: Volume taken from sample to be derivatizized (Sample_Volume_derivatized) -> divide by 1000 to get mL
* $V_4$: Volume TMSH used for derivatization(TMSH_Volume) -> divide by 1000 to get mL
* $V_5$: Volume TMSH + Volume sample ($V_3$ + $V_4$)
* $c_3$: Concentration internal standard measured -> c_si_measured
* $c_2$: Concentration internal standard after sample dissolved -> c_si_dissolved -> microgram per mL

$$ c_3 = {\frac{c_2 \cdot V_3} {V_5}} $$
$$ c_3 = {\frac{c_2 \cdot V_3} {V_3 + V_4}} $$
```{r, message=FALSE, warning=FALSE}
recovery$c_si_measured <- with(recovery,
                               ((c_si_dissolved*(Sample_Volume_derivatized/1000))/
                              ((TMSH_Volume/1000)+(Sample_Volume_derivatized/1000))))
```

Calculate recovery

$$ recovery = {\frac{concentration_{measured}} {concentration_{internal \ standard} }} $$

```{r, message=FALSE, warning=FALSE}
recovery$rec <- with(recovery, conc_measured_LOQ/c_si_measured)
```

Get recovery of 23:0, since 17:0 mostly below LOQ or not detected at all
```{r, message=FALSE, warning=FALSE}
recovery <- filter(recovery, name_pufa == "23:0")
```

Calculate mean recovery for Jan-Mar, May-Jun and Nov
```{r, message=FALSE, warning=FALSE}
recovery <- ddply(recovery, c("cali", "cali_pufa"), summarise, mean_rec = mean(rec, na.rm = TRUE))
```

# Concentration in microcosm samples 
## Prepare data

Merge microcosm and calibration data
```{r, message=FALSE, warning=FALSE}
microcosm$cali <-with(microcosm, ifelse(Date > "2020-01-01" & Date <= "2020-04-01","cali_1",ifelse(Date > "2020-05-01" & Date <= "2020-07-01","cali_2","cali_3")))

unique(microcosm$cali)

filter(microcosm, is.na(microcosm$cali)) #NA: "chiro fishfood 2 31.05.2019" lost during pipetting, test chiro A and B -> so real NA 

microcosm <- filter(microcosm, !is.na(cali)) # remove NA
```

column with info needed to merge correctly
```{r, message=FALSE, warning=FALSE}
microcosm <- unite(microcosm, "cali_pufa", "cali", "name_pufa", sep = "_", remove = FALSE)
```
colums needed from standards to calculate conc. in samples
```{r, message=FALSE, warning=FALSE}
cali <- select(calibration, c("cali_pufa", "slope")) 
```

for every standard measurement in same calibration curve and substance slope, which are the same
```{r, message=FALSE, warning=FALSE}
cali <- distinct(cali) 
```

Merge calibration and microcosm
```{r, message=FALSE, warning=FALSE}
microcosm <- left_join(microcosm, cali, by = "cali_pufa")
```

Merge with LOQ table
```{r, message=FALSE, warning=FALSE}
microcosm <- left_join(microcosm, LOQ, by = "cali_pufa")

filter(microcosm, is.na(min_c2)) # NAs okay, because this PUFA not identified in Nov., identification started at C16

microcosm <- filter(microcosm, !is.na(min_c2)) # remove NAs
```

# Calculate concentration in samples
## PUFA dissolved
Measured concentration in microgramm per mL

$$ area = slope \cdot concentration $$
$$ concentration = {\frac  {area } {slope}} $$

```{r, message=FALSE, warning=FALSE}
microcosm$conc_measured <- with(microcosm, (area)/slope)
```

## Include LOQ

Set 0 for concentration < LOQ 
```{r, message=FALSE, warning=FALSE}
microcosm$conc_measured_LOQ <- with(microcosm, ifelse(conc_measured < min_c2, 0, conc_measured))
```
## Include recovery
Chose columns needed 
```{r, message=FALSE, warning=FALSE}
mean_rec <- select(recovery, -c("cali_pufa"))
```

Merge microcosm and recovery data
```{r, message=FALSE, warning=FALSE}
microcosm <- left_join(microcosm, mean_rec, by = c("cali"))
```

Include recovery in concentration
```{r, message=FALSE, warning=FALSE}
microcosm$conc_measured_LOQ_rec <- with(microcosm, conc_measured_LOQ/mean_rec)
```

## Concentration in sample
Get concentration dissolved in MeOH after evaporation
* $V_2$: Volume MeOH used to dissolve sample (MeOH_Volume_dissolved) -> divide by 1000 to get mL
* $V_3$: Volume taken from sample to be derivatizized (Sample_Volume_derivatized) -> divide by 1000 to get mL
* $V_4$: Volume TMSH used for derivatization(TMSH_Volume) -> divide by 1000 to get mL
* $V_5$: Volume TMSH + Volume sample
* $c_3$: Concentration measured (including corrections with LOQ and recovery) -> conc_measured_LOQ_rec
* $c_2$: Concentration after sample dissolved -> c_dissolved -> microgram per mL

-> Calculate c2:

$$ c_2 = {\frac{c_3 \cdot V_5} {V_2}} $$
$$ c_2 = {\frac{c_3 \cdot (V_3 + V_4)} {V_2}} $$
```{r, message=FALSE, warning=FALSE}
microcosm$c_dissolved <-with(microcosm,
                            (conc_measured_LOQ_rec*((TMSH_Volume/1000)+
                              (Sample_Volume_derivatized/1000)))/
                                MeOH_Volume_dissolved)
```
Mass PUFA in dissolved sample
* $V_2$: Volume MeOH used to dissolve sample (MeOH_Volume_dissolved) -> divide by 1000 to get mL
* $c_2$: Concentration after sample dissolved -> c_dissolved -> microgram per mL
* m: mass PUFA dissolved in MeOH

$$ m = c_2 \cdot V_2 $$
```{r, message=FALSE, warning=FALSE}
microcosm$mass_pufa <- with(microcosm, c_dissolved*(MeOH_Volume_dissolved/1000))

range(microcosm$mass_pufa) #-> 0.00000 15.79341 microgramm

range(microcosm$dry_weight_mg) # 0.4239 40.8680 mg
```
Weight PUFA per dry weight of organism (microgramm PUFA per milligram dry mass)

$$ m_{dry.pufa} = {\frac{mass_{pufa}} {dry_{weight}}} $$

```{r, message=FALSE, warning=FALSE}
microcosm$dry_pufa <- with(microcosm, mass_pufa/dry_weight_mg)
```



# microcosm data for statistics

Merge microcosm complete and microcosm
```{r, message=FALSE, warning=FALSE}
microcosm <- left_join(microcosm, microcosm_complete, by = c( "experiment", "GC_ID", "dry_weight_mg", "Internal_Standard_Volume_ID", "MeOH_Volume_dissolved", "TMSH_Volume", "Sample_Volume_derivatized", "runinfo", "area", "name_pufa", "runname","Date"))
```
Chose needed columns
```{r, message=FALSE, warning=FALSE}
microcosm <- select(microcosm, c("Organism", "treatment", "experiment", "experiment_ID", "name_pufa", "runinfo", "runinfo_original","runname", "dry_pufa", "dry_weight_mg"))
```
Chose PUFA for statistics
```{r, message=FALSE, warning=FALSE}
name_stats <- c("24:1n-9", "24:0", "22:6n-3", "22:2n-6", "22:1n-9", "22:0", "20:5n-3", "20:3n-3", "20:4n-6", "20:3n-6", "20:2n-6", "20:1n-9", "20:0", "18:3n-3", "18:3n-6", "18:2n-6c", "18:2n-6t", "18:1n-7", "18:1n-9c", "18:1n-9t", "18:0")

microcosm <- filter(microcosm, name_pufa %in% name_stats)
```
# Save data
```{r, message=FALSE, warning=FALSE}
#write.csv(microcosm, file = "microcosm_pufa.csv", row.names = FALSE)
```
